---
title: "3_quality_check_filtering_treated_worms"
author: "Hailey Sounart"
date: "2/28/2023"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(ggplot2)
  library(ggpubr)
```

```{r set working directory}

setwd("<path to working directory>")

``` 

# Load pre-filtered object, contains metadata

## Read in Object

``` {r read in labeled object}

all_section_QC <- readRDS("bmalayi.treated.pre.filtered.rds")
all_section_QC
```

# Remove Wolbachia (wBm) from count matrix

``` {r get wBm gene list}

#load in wBm genes
wBm_all_gene_list <- readRDS("wBm_all_genes/all_wBm_genes.rds") #refhook = NULL
length(wBm_all_gene_list)

all_section_QC_wBm <- intersect(rownames(all_section_QC), wBm_all_gene_list)
all_section_QC_wBm

all_section_QC_wBm_obj <- SubsetSTData(all_section_QC, features = all_section_QC_wBm)
all_section_QC_wBm_obj

```

``` {r get bmalayi genes}

#load in wBm genes
bmalayi_all_gene_list <- readRDS("bmalayi_all_genes/bmalayi_genes.rds")
length(bmalayi_all_gene_list)

all_section_QC_bmalayi <- intersect(rownames(all_section_QC), bmalayi_all_gene_list)
length(all_section_QC_bmalayi)

all_section_QC_bmalayi_obj <- SubsetSTData(all_section_QC, features = all_section_QC_bmalayi)
all_section_QC_bmalayi_obj


```

# Remove genes with 0 counts

## Brugia
``` {r remove genes}

#min.gene.count
nUMI_filt = Matrix::rowSums(all_section_QC_bmalayi_obj@assays$RNA@counts)
gene_umi_count <- rownames(all_section_QC_bmalayi_obj@assays$RNA[(nUMI_filt >= 1),])

all_section_QC_bmalayi_obj.gene.filt <- SubsetSTData(all_section_QC_bmalayi_obj, features = gene_umi_count)

cat("Genes removed: ", nrow(all_section_QC_bmalayi_obj) - nrow(all_section_QC_bmalayi_obj.gene.filt), "\n")
cat("Genes remaining: ", nrow(all_section_QC_bmalayi_obj.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(all_section_QC_bmalayi_obj.gene.filt)/nrow(all_section_QC_bmalayi_obj))*100), "\n")

all_section_QC_bmalayi_obj.gene.filt

```

Check metadata
``` {r check metadata}

all_section_QC_bmalayi_obj.gene.filt@meta.data

```

## wBm
``` {r remove filtered genes}

#min.gene.count
nUMI_filt = Matrix::rowSums(all_section_QC_wBm_obj@assays$RNA@counts)
gene_umi_count <- rownames(all_section_QC_wBm_obj@assays$RNA[(nUMI_filt >= 1),])

all_section_QC_wBm_obj.gene.filt <- SubsetSTData(all_section_QC_wBm_obj, features = gene_umi_count)

cat("Genes removed: ", nrow(all_section_QC_wBm_obj) - nrow(all_section_QC_wBm_obj.gene.filt), "\n")
cat("Genes remaining: ", nrow(all_section_QC_wBm_obj.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(all_section_QC_wBm_obj.gene.filt)/nrow(all_section_QC_wBm_obj))*100), "\n")

all_section_QC_wBm_obj.gene.filt

```

Check metadata
``` {r check metadata again}

all_section_QC_wBm_obj.gene.filt@meta.data

```

# Save >0 counts gene object

``` {r save labeled object}

saveRDS(all_section_QC_bmalayi_obj.gene.filt, file = "bmalayi.treated.0.gene.filt.rds")

saveRDS(all_section_QC_wBm_obj.gene.filt, file = "wBm.treated.0.gene.filt.rds")

```


# QC plots - Brugia 

``` {r read in labeled object}

#Read in object
bmalayi.treated.0.gene.filt <- readRDS("bmalayi.treated.0.gene.filt.rds")
bmalayi.treated.0.gene.filt

```

## Add wolbachia metadata to brugia object

```{r filter all gene object}

bmalayi.treated.0.gene.filt@meta.data$wolbachia <- ifelse((bmalayi.treated.0.gene.filt@meta.data$counts.wBm >= 1), 1, 0)
bmalayi.treated.0.gene.filt@meta.data

```

## Split into control and doxy

``` {r split objs}

length(bmalayi.treated.0.gene.filt@meta.data$condition[(bmalayi.treated.0.gene.filt@meta.data$condition == "control")])
length(bmalayi.treated.0.gene.filt@meta.data$condition[(bmalayi.treated.0.gene.filt@meta.data$condition == "doxycycline")])
```

``` {r split objs 2}

ctrl_spots <- SubsetSTData(bmalayi.treated.0.gene.filt, expression = condition == "control")
ctrl_spots
doxy_spots <- SubsetSTData(bmalayi.treated.0.gene.filt, expression = condition == "doxycycline")
doxy_spots
```

## Statistics

Average UMIs and unique genes per spot
``` {r get stats}

mean(ctrl_spots@meta.data$nCount_RNA)
mean(ctrl_spots@meta.data$nFeature_RNA)

mean(doxy_spots@meta.data$nCount_RNA)
mean(doxy_spots@meta.data$nFeature_RNA)

```
Doxycycline treated worms: average of 295 UMIs per spot and 163 unique genes per spot

Total genes detected in treated worms (remove genes with 0 counts in doxycycline subset count matrix)
``` {r remove genes}

#min.gene.count
nUMI_filt = Matrix::rowSums(doxy_spots@assays$RNA@counts)
gene_umi_count <- rownames(doxy_spots@assays$RNA[(nUMI_filt >= 1),])

doxy_spots.gene.filt <- SubsetSTData(doxy_spots, features = gene_umi_count)

cat("Genes removed: ", nrow(doxy_spots) - nrow(doxy_spots.gene.filt), "\n")
cat("Genes remaining: ", nrow(doxy_spots.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(doxy_spots.gene.filt)/nrow(doxy_spots))*100), "\n")

doxy_spots.gene.filt

```
2679 total genes detected in doxycycline treated worm sections

Spots with Wolbachia
``` {r split objs by wBm}

doxy_spots@meta.data[(doxy_spots@meta.data$wolbachia == "1"), ]

doxy_spots@meta.data[(doxy_spots@meta.data$wolbachia == "0"), ]

```
18 spots with Wolbachia, 98 spots without Wolbachia (pre-filtering) (15.5%)

## Violin plots - per condition

```{r prep cond violins}

cond_levels <- c("control","doxycycline")
bmalayi.treated.0.gene.filt@meta.data$condition <- factor(x = bmalayi.treated.0.gene.filt@meta.data$condition, levels = cond_levels)

``` 


```{r plot cond violins}

# Figure 6D

fill <- c("#e78ac3","#fc8d62")

bmalayi_vlnplot_unique_gene_cond_no_pt <- VlnPlot(bmalayi.treated.0.gene.filt, features = c("nFeature_RNA"), split.by = "condition", group.by = "wolbachia", cols = fill, pt.size = 0)
bmalayi_vlnplot_umi_cond_no_pt  <- VlnPlot(bmalayi.treated.0.gene.filt, features = c("nCount_RNA"), split.by = "condition", group.by = "wolbachia", cols = fill, pt.size = 0)

``` 


```{r plot cond violins 2}

# Figure 6D

bmalayi_vlnplot_unique_gene_cond_no_pt 
bmalayi_vlnplot_umi_cond_no_pt  

``` 

```{r}
dir.create(paste0("qc_plots"))
```

```{r plot cond violins 3}

# Figure 6D

ggsave(file = "Figure_6D_umi.pdf", plot = bmalayi_vlnplot_umi_cond_no_pt, dpi = 300)
ggsave(file = "Figure_6D_genes.pdf", plot = bmalayi_vlnplot_unique_gene_cond_no_pt, dpi = 300)

``` 

## Violin plots - per section for treated worms

```{r find unique sections}
unique(doxy_spots@meta.data$sample_section)
length(unique(doxy_spots@meta.data$sample_section))

``` 

```{r unique sections 2}

section_levels <- c("T1_1","T1_3","T1_4","T1_5","T2_1")
doxy_spots@meta.data$sample_section <- factor(x = doxy_spots@meta.data$sample_section, levels = section_levels)

``` 

```{r plot unique sections}

# Figure 6A

treated_sects_fill <- c("#8da0cb", "#8da0cb", "#8da0cb", "#8da0cb", "#8da0cb")

bmalayi.treated_vlnplot_unique_gene_section_no_pt <- VlnPlot(doxy_spots, features = c("nFeature_RNA"), split.by = "sample_section", group.by = "sample_section", cols = treated_sects_fill, pt.size = 0)
bmalayi.treated_vlnplot_umi_section_no_pt  <- VlnPlot(doxy_spots, features = c("nCount_RNA"), split.by = "sample_section", group.by = "sample_section", cols = treated_sects_fill, pt.size = 0) 

bmalayi.treated_vlnplot_unique_gene_section_no_pt
bmalayi.treated_vlnplot_umi_section_no_pt

``` 
```{r plot unique sections 2}

ggsave(file = "Figure_6A_umi.pdf", plot = bmalayi.treated_vlnplot_umi_section_no_pt, dpi = 300)
ggsave(file = "Figure_6A_gene.pdf", plot = bmalayi.treated_vlnplot_unique_gene_section_no_pt, dpi = 300)

``` 

## Correlation plots - per sample for treated worms

```{r make treated worms smple matrix}


umis.all.avg <- doxy_spots@assays$RNA@counts
sum.expr.all.avg <- lapply(unique(doxy_spots$worm_sample), function(treat) {
  umis_subset.all.avg <- umis.all.avg[, doxy_spots$worm_sample %in% treat]  #subset counts dataframe
  print(class(umis_subset.all.avg))
  log1p(Matrix::rowMeans(umis_subset.all.avg))
})

sum.expr.all.avg <- as.data.frame(sum.expr.all.avg)
colnames(sum.expr.all.avg) <- unique(doxy_spots$worm_sample)
worm_sample_raw_unfil_df <- sum.expr.all.avg
worm_sample_raw_unfil_df
```

Treated worm sample comparisons
```{r treated worms smpl corrs}

colnames(worm_sample_raw_unfil_df) <- c("T1", "T2")
worm_sample_raw_unfil_df

p1 <- ggplot(worm_sample_raw_unfil_df, aes(T1, T2)) + 
  geom_point(alpha = 1,colour = "#8da0cb") + 
  stat_cor(method = "pearson", label.x = 0, label.y = 3.7, p.accuracy = 1e-323) + 
  geom_smooth(method = "lm", fullrange=TRUE, color="#8da0cb", se = TRUE, level = 0.95)  + 
  theme_minimal() + 
  labs(title="Average gene expression between treated samples",
       x="Treated worm T1 (log (1 + avg raw UMI counts per spot))", 
       y = "Treated worm T2 (log (1 + avg raw UMI counts per spot))") +
  theme(axis.text.y = element_text(hjust = 1, color = "black"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5),
        panel.background = element_blank(),  # Remove panel background
        axis.ticks.length = unit(0.25, "cm"),  # Adjust the length of the tick marks if needed
        axis.text.x = element_text(color = "black"))

p1

ggsave(file = "qc_plots/bmalayi.treated_corr_worm_smpl.pdf", plot = p1, dpi = 300)

``` 

## Correlation plots - per section for treated worms

```{r treated worms section matrix}


umis.all.avg <- doxy_spots@assays$RNA@counts
sum.expr.all.avg <- lapply(unique(doxy_spots$sample_section), function(treat) {
  umis_subset.all.avg <- umis.all.avg[, doxy_spots$sample_section %in% treat]  #subset counts dataframe
  print(class(umis_subset.all.avg))
  log1p(Matrix::rowMeans(umis_subset.all.avg))
})

sum.expr.all.avg <- as.data.frame(sum.expr.all.avg)
colnames(sum.expr.all.avg) <- unique(doxy_spots$sample_section)
worm_section_raw_unfil_df <- sum.expr.all.avg
worm_section_raw_unfil_df
```

Treated worm section comparisons
```{r treated worms section corrs}

# Figure 6B plots

for (i in colnames(worm_section_raw_unfil_df)) {
  first_section <- i
  first_section <- noquote(first_section)
  print(first_section)

  for (j in colnames(worm_section_raw_unfil_df)){ 
    
    second_section <- noquote(j)
    print(second_section)
    
    p <- aes_(as.name(first_section), as.name(second_section))
    print(p)
    
    p <- ggplot(worm_section_raw_unfil_df, aes_(as.name(first_section),as.name(second_section))) + 
      geom_point(alpha = 1,colour = "#8da0cb") + 
      stat_cor(method = "pearson", label.x = 0, label.y = 3.7, p.accuracy = 1e-323) + 
      geom_smooth(method = "lm", fullrange=TRUE, color="#8da0cb", se = TRUE, level = 0.95)  + 
      theme_minimal() + 
      labs(title="Average gene expression between treated samples", 
           x=paste0("Treated worm ", first_section, "(log (1 + avg raw UMI counts per spot))"), 
           y = paste0("Treated worm", second_section, "(log (1 + avg raw UMI counts per spot))")) +
      theme(axis.text.y = element_text(hjust = 1, color = "black"),
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.ticks = element_line(color = "black"),
            axis.line = element_line(color = "black"),
            plot.title = element_text(hjust = 0.5),
            panel.background = element_blank(),  # Remove panel background
            axis.ticks.length = unit(0.25, "cm"),  # Adjust the length of the tick marks if needed
            axis.text.x = element_text(color = "black"))
    
    ggsave(file = paste0("qc_plots/Figure_6B_", first_section, "_", second_section, ".pdf"), plot = p, dpi = 300)
  }
}

``` 

# QC plots - Wolbachia

``` {r read in wBm filt object}

#Read in object
wBm.treated.0.gene.filt <- readRDS("wBm.treated.0.gene.filt.rds")
wBm.treated.0.gene.filt

```

## Violin plots - per condition

```{r wBm violin plots overall}

cond_levels <- c("control","doxycycline")
wBm.treated.0.gene.filt@meta.data$condition <- factor(x = wBm.treated.0.gene.filt@meta.data$condition, levels = cond_levels)

``` 


```{r wBm violin plots overall 2}

wBm_vln_fill <- c("#66c2a5","#8da0cb")

wBm_vlnplot_unique_gene_cond_no_pt <- VlnPlot(wBm.treated.0.gene.filt, features = c("nFeature_RNA"), split.by = "condition", group.by = "condition", cols = wBm_vln_fill, pt.size = 0, log = TRUE)
wBm_vlnplot_umi_cond_no_pt  <- VlnPlot(wBm.treated.0.gene.filt, features = c("nCount_RNA"), split.by = "condition", group.by = "condition", cols = wBm_vln_fill, pt.size = 0, log = TRUE)

``` 


```{r wBm violin plots overall 3}

wBm_vlnplot_unique_gene_cond_no_pt 
wBm_vlnplot_umi_cond_no_pt  

``` 

```{r violin plots sample 4}

# Figure 6C

ggsave(file = "Figure_6C_gene.pdf", plot = wBm_vlnplot_unique_gene_cond_no_pt , dpi = 300)
ggsave(file = "Figure_6C_umi.pdf", plot = wBm_vlnplot_umi_cond_no_pt, dpi = 300)

``` 

# Load protein coding genes, no ribo, no mito
```{r load protein coding genes}

bmalayi_protein_coding.no.ribo.no.mito <- readRDS("bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")
length(bmalayi_protein_coding.no.ribo.no.mito)

``` 

# Keep only protein coding genes (no mito and no ribo)

```{r keep only protein coding genes}

bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito <- SubsetSTData(bmalayi.treated.0.gene.filt, features = rownames(bmalayi.treated.0.gene.filt)[(rownames(bmalayi.treated.0.gene.filt) %in% bmalayi_protein_coding.no.ribo.no.mito)])
bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito

``` 

# Save partially filtered bmalayi objects

``` {r save pre filtered treated bmalayi object}

saveRDS(bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito, file = "bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito.rds")

```

# Filtering Part 2

## Split into control & doxy objects

``` {r split objs once again}

ctrl_spots_filt <- SubsetSTData(bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito, expression = condition == "control")
ctrl_spots_filt
doxy_spots_filt <- SubsetSTData(bmalayi.treated.0.gene.filt.protein.coding.no.ribo.no.mito, expression = condition == "doxycycline")
doxy_spots_filt
```

## Filter control spots 

Filter low quality spots
```{r filter ctrl spots}

#same filtering criteria as used previously for the control worms
ctrl_spots.spot.filt <- SubsetSTData(ctrl_spots_filt, expression = nFeature_RNA >= 30 & nCount_RNA >= 50 & percent.mito < 3 & percent.ribo < 30 & nCount_RNA < 10000)
cat("Spots removed: ", ncol(ctrl_spots_filt) - ncol(ctrl_spots.spot.filt), "\n")
cat("Spots remaining: ", ncol(ctrl_spots.spot.filt), "\n")
cat("Percent Spots remaining: ", (ncol(ctrl_spots.spot.filt)/ncol(ctrl_spots_filt)*100), "\n")

ctrl_spots.spot.filt

```

## Filter treated spots 

Filter low quality spots
```{r filter treated spots}

#different filtering for treated worms, due to their differing count distributions
doxy_spots.spot.filt <- SubsetSTData(doxy_spots_filt, expression = nFeature_RNA >= 10 & nCount_RNA >= 15 & percent.mito < 5 & percent.ribo < 35)
cat("Spots removed: ", ncol(doxy_spots_filt) - ncol(doxy_spots.spot.filt), "\n")
cat("Spots remaining: ", ncol(doxy_spots.spot.filt), "\n")
cat("Percent Spots remaining: ", (ncol(doxy_spots.spot.filt)/ncol(doxy_spots_filt)*100), "\n")

doxy_spots.spot.filt

```

## Merge conditions back together

```{r filter all gene object}

ctrl_doxy_spot_filt_merged <- MergeSTData(x = ctrl_spots.spot.filt, y = doxy_spots.spot.filt)
ctrl_doxy_spot_filt_merged
ctrl_doxy_spot_filt_merged@meta.data

```

## Filter genes

Filter based on number of spots per gene & total UMI counts per gene.
```{r filter genes}

#min.gene.spots
nSpots_filt = Matrix::rowSums(ctrl_doxy_spot_filt_merged@assays$RNA@counts > 0)
gene_spot_count <- rownames(ctrl_doxy_spot_filt_merged@assays$RNA[(nSpots_filt >= 1),])
  
#min.gene.count
nUMI_filt = Matrix::rowSums(ctrl_doxy_spot_filt_merged@assays$RNA@counts)
gene_umi_count <- rownames(ctrl_doxy_spot_filt_merged@assays$RNA[(nUMI_filt >= 1),])

ctrl_doxy_spot_filt_merged.spot.gene.filt <- SubsetSTData(ctrl_doxy_spot_filt_merged, features = gene_umi_count)
ctrl_doxy_spot_filt_merged.spot.gene.filt <-SubsetSTData(ctrl_doxy_spot_filt_merged.spot.gene.filt, features = gene_spot_count)

cat("Genes removed: ", nrow(ctrl_doxy_spot_filt_merged) - nrow(ctrl_doxy_spot_filt_merged.spot.gene.filt), "\n")
cat("Genes remaining: ", nrow(ctrl_doxy_spot_filt_merged.spot.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(ctrl_doxy_spot_filt_merged.spot.gene.filt)/nrow(ctrl_doxy_spot_filt_merged))*100), "\n")


ctrl_doxy_spot_filt_merged.spot.gene.filt

```

## Re-calc metadata

```{r recalc metadata load gene lists}

#load in mitochondrial gene list
bmalayi_mito_gene_list <- readRDS("mitochondrial_genes/bmalayi_mito_genes.rds")
bmalayi_mito_gene_list

#load in rRNA genes
bmalayi_rRNA_gene_list <- readRDS("rRNA_genes/bmalayi_rRNA_genes.rds")
bmalayi_rRNA_gene_list

#load in ribosomal protein coding genes
bmalayi_ribo_gene_list <- readRDS("ribosomal_genes/bmalayi_ribo_genes.rds")
#bmalayi_ribo_gene_list

#combine rRNA and ribosomal protein coding genes into a single list
bmalayi_all_ribo_list <- c(bmalayi_rRNA_gene_list, bmalayi_ribo_gene_list)
bmalayi_all_ribo_list

#load in bmalayi protein coding 
bmalayi_coding_list <- readRDS("all_protein_coding_genes/bmalayi_protein_coding_genes.rds")

#load in bmalayi protein coding, no ribo
bmalayi_coding_no_ribo_list <- readRDS("bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.rds")

#load in bmalayi protein coding, no mito
bmalayi_coding_no_mito_list <- readRDS("bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.rds")

#load in bmalayi protein coding,no ribo and no mito
bmalayi_coding_no_ribo_no_mito_list <- readRDS("bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")

```

```{r recalc metadata}

print(sum(ctrl_doxy_spot_filt_merged.spot.gene.filt$percent.mito))
print(sum(ctrl_doxy_spot_filt_merged.spot.gene.filt$counts.mito))


ctrl_doxy_spot_filt_merged.spot.gene.filt.new <- ctrl_doxy_spot_filt_merged.spot.gene.filt

mito_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_mito_gene_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.mito <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[mito_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.mito <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[mito_in_obj, ])
rRNA_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_rRNA_gene_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.rRNA <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[rRNA_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.rRNA <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[rRNA_in_obj, ])
ribo_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_all_ribo_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.ribo <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[ribo_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.ribo <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[ribo_in_obj, ])
coding_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_coding_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.protein_coding <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[coding_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.protein_coding <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[coding_in_obj, ])
coding_no_ribo_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_coding_no_ribo_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.protein_coding_no_ribo <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[ coding_no_ribo_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.protein_coding_no_ribo <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[ coding_no_ribo_in_obj, ])
coding_no_mito_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_coding_no_mito_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.protein_coding_no_mito <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[coding_no_mito_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.protein_coding_no_mito <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[coding_no_mito_in_obj, ])
coding_no_ribo_no_mito_in_obj <- Reduce(intersect, list(rownames(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA),bmalayi_coding_no_ribo_no_mito_list))
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.protein_coding_no_ribo_no_mito <- (Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts[coding_no_ribo_no_mito_in_obj, ])/Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt@assays$RNA@counts))*100
ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.protein_coding_no_ribo_no_mito <- Matrix::colSums(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA@counts[coding_no_ribo_no_mito_in_obj, ])


print(sum(ctrl_doxy_spot_filt_merged.spot.gene.filt.new$percent.mito))
print(sum(ctrl_doxy_spot_filt_merged.spot.gene.filt.new$counts.mito))

ctrl_doxy_spot_filt_merged.spot.gene.filt.new

```

## Save spot, gene filt objects
``` {r save filtered bmalayi object}

saveRDS(ctrl_doxy_spot_filt_merged.spot.gene.filt.new, file = "bmalayi.ctrl.doxy.filt.obj.rds")

```

## Check Wolbachia presence

### Split into control and doxy

``` {r split objs again again}

length(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@meta.data$condition[(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@meta.data$condition == "control")])
length(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@meta.data$condition[(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@meta.data$condition == "doxycycline")])
```

``` {r split objs again 2}

ctrl_spots <- SubsetSTData(ctrl_doxy_spot_filt_merged.spot.gene.filt.new, expression = condition == "control")
ctrl_spots
doxy_spots <- SubsetSTData(ctrl_doxy_spot_filt_merged.spot.gene.filt.new, expression = condition == "doxycycline")
doxy_spots
```

Spots with Wolbachia
``` {r split treated by wBm}

doxy_spots@meta.data[(doxy_spots@meta.data$wolbachia == "1"), ]

doxy_spots@meta.data[(doxy_spots@meta.data$wolbachia == "0"), ]

```
18 out of 97 spots are Wolbachia+ in doxy treated worms (18.6%)

# Wolbachia processing

``` {r read in wBm treated filt obj}

#Read in object
wBm.treated.0.gene.filt <- readRDS("wBm.treated.0.gene.filt.rds")
wBm.treated.0.gene.filt

```

Subset Wolbachia object to only include those spots in the brugia object

```{r filter wBm obj}

unique(wBm.treated.0.gene.filt$worm_sample)

bmalayi.treated_obj.spot.gene.filt.spots <- colnames(ctrl_doxy_spot_filt_merged.spot.gene.filt.new@assays$RNA[,])
length(bmalayi.treated_obj.spot.gene.filt.spots)

wBm.treated.0.gene.filt.brugia.filt.spots <- SubsetSTData(wBm.treated.0.gene.filt, spots = bmalayi.treated_obj.spot.gene.filt.spots)

wBm.treated.0.gene.filt.brugia.filt.spots

```

Remove genes with 0 summed UMIs - from R objects
```{r remove 0 count genes}

#min.gene.spots
nSpots_filt = Matrix::rowSums(wBm.treated.0.gene.filt.brugia.filt.spots@assays$RNA@counts > 0)
gene_spot_count <- rownames(wBm.treated.0.gene.filt.brugia.filt.spots@assays$RNA[(nSpots_filt >= 1),])
  
#min.gene.count
nUMI_filt = Matrix::rowSums(wBm.treated.0.gene.filt.brugia.filt.spots@assays$RNA@counts)
gene_umi_count <- rownames(wBm.treated.0.gene.filt.brugia.filt.spots@assays$RNA[(nUMI_filt >= 1),])

wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt <- SubsetSTData(wBm.treated.0.gene.filt.brugia.filt.spots, features = gene_umi_count)
wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt <-SubsetSTData(wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt, features = gene_spot_count)

cat("Genes removed: ", nrow(wBm.treated.0.gene.filt.brugia.filt.spots) - nrow(wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt), "\n")
cat("Genes remaining: ", nrow(wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt)/nrow(wBm.treated.0.gene.filt.brugia.filt.spots))*100), "\n")


wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt


``` 

## Save Wolbachia objects

```{r save wBm obj}

saveRDS(wBm.treated.0.gene.filt.brugia.filt.spots.gene.filt, file = "wBm.treated.filt.rds")

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

Outputs are the violin plots and correlation plots shown in Figure 6A-D of the main manuscript, the filtered brugia count matrix and metadata rds object used in the next script, 4_normalization_treated_worms.Rmd, and the Wolbachia (wBm) filtered count matrix and metadata rds object.
