---
title: "4_Normalization"
author: "Hailey Sounart"
date: "3/1/2023"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
```

```{r set working directory}

setwd("<path to working directory>")

``` 

# Read in Filtered Seurat Objects

Read in filtered object
``` {r read filt object}

bmalay.treated_filt_obj <- readRDS("bmalayi.ctrl.doxy.filt.obj.rds")
bmalay.treated_filt_obj

```

# Normalization

## Split per section

```{r subset each worm section}

ctrl_doxy_filt_obj.unique.sect.names <- unique(bmalay.treated_filt_obj$sample_section)
ctrl_doxy_filt_obj.unique.sect.names

ctrl_doxy_filt_obj.section.list.nfeat <- lapply(ctrl_doxy_filt_obj.unique.sect.names, function(i){SubsetSTData(bmalay.treated_filt_obj, expression = (sample_section == i))})

ctrl_doxy_filt_obj.section.list.nfeat
```

## Normalize each section object separately

Unequal sections per sample for treated worms compared to control worms, will normalize each section separately.
```{r normalize each section object separately}

for (i in 1:length(ctrl_doxy_filt_obj.section.list.nfeat)) {
  ctrl_doxy_filt_obj.section.list.nfeat[[i]] <- SCTransform(ctrl_doxy_filt_obj.section.list.nfeat[[i]], vars.to.regress = "nFeature_RNA", return.only.var.genes = FALSE)
}

```

## Data integration
```{r data intregration}

integration.features<- SelectIntegrationFeatures(object.list = ctrl_doxy_filt_obj.section.list.nfeat, nfeatures = 7000)

#modifying the maximum size of global variables
options(future.globals.maxSize = 8000 * 1024^8)

ctrl_doxy_filt_obj.section.list.nfeat.prep <- PrepSCTIntegration(object.list = ctrl_doxy_filt_obj.section.list.nfeat, anchor.features = integration.features, verbose = FALSE)

```

```{r save integration feats}

saveRDS(integration.features, file = "integration.features.treated.rds")

```

## Merge

``` {r merge normalized object}

ctrl_doxy_filt_obj.section.list.nfeat.prep.merged <- MergeSTData(x = ctrl_doxy_filt_obj.section.list.nfeat.prep[[1]], y = ctrl_doxy_filt_obj.section.list.nfeat.prep[2:length(ctrl_doxy_filt_obj.section.list.nfeat.prep)], merge.data = TRUE)

ctrl_doxy_filt_obj.section.list.nfeat.prep.merged

```

## Save Normalized object

```{r save normalized object}

#save control,doxy normalized data, merged
ctrl_doxy_filt_obj.section.list.nfeat.prep.merged

saveRDS(ctrl_doxy_filt_obj.section.list.nfeat.prep.merged, file = "bmalay.treated.normalized.obj.rds")

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

The outputs are the integration features from SelectIntegrationFeatures, output of PrepSCTIntegration, and the normalized rds object of brugia count matrix and metadata. The normalized rds object of brugia count matrix and metadata are used in the scripts 5_treated_worm_analysis_r_linux.Rmd, 6_look_at_treated_analysis.Rmd, and 7_genes_on_tissue_visualization_treated_worms.Rmd.
