---
title: "7_genes_on_tissue_visualization_treated_worms"
author: "Hailey Sounart"
date: "5/21/2021"
output: html_document
---


````{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(magrittr)
  library(ggplot2)
  library(magick)
  library(grid)
```

```{r set working directory}

#setwd("<path to working directory>")
setwd("/Users/hailey.sounart/Documents/brugia_malayi/230911_final_code/r_scripts_treated_worms")
``` 


# Read in Normalized object

``` {r read norm object}

bmalay.treated_norm_obj <- readRDS("bmalay.treated.normalized.obj.rds")
bmalay.treated_norm_obj 

```

# Add wolbachia metadata to brugia object

```{r wol labels}

bmalay.treated_norm_obj@meta.data$wolbachia <- ifelse((bmalay.treated_norm_obj@meta.data$counts.wBm >= 1), 1, 0)
bmalay.treated_norm_obj@meta.data

```

# Only for doxy worms

## Split into control and doxy

``` {r split objs}

length(bmalay.treated_norm_obj@meta.data$condition[(bmalay.treated_norm_obj@meta.data$condition == "control")])
length(bmalay.treated_norm_obj@meta.data$condition[(bmalay.treated_norm_obj@meta.data$condition == "doxycycline")])
```

``` {r split objs 2}

ctrl_spots <- SubsetSTData(bmalay.treated_norm_obj, expression = condition == "control")
ctrl_spots
doxy_spots <- SubsetSTData(bmalay.treated_norm_obj, expression = condition == "doxycycline")
doxy_spots
```

## Load raw image
```{r load raw image}

# Path to the raw HE images used as input for spaceranger

high_res_image_list <- c("/Users/hailey.sounart/Documents/brugia_malayi/old_analysis/all_sections_analysis//high_res_images/210304_V10T03_275_Worms_HE_HS-Spot000001_25perc.jpg",
                                "/Users/hailey.sounart/Documents/brugia_malayi/old_analysis/all_sections_analysis//high_res_images/210304_V10T03_275_Worms_HE_HS-Spot000001_25perc.jpg",
                                "/Users/hailey.sounart/Documents/brugia_malayi/old_analysis/all_sections_analysis//high_res_images/210304_V10T03_275_Worms_HE_HS-Spot000001_25perc.jpg",
                                "/Users/hailey.sounart/Documents/brugia_malayi/old_analysis/all_sections_analysis//high_res_images/210304_V10T03_275_Worms_HE_HS-Spot000001_25perc.jpg",
                                "/Users/hailey.sounart/Documents/brugia_malayi/old_analysis/all_sections_analysis//high_res_images/210208_worm_bmalayi_V10T03_300_HE_HS_B1-Spot000001_25perc.jpg")

```

## Load the images

```{r load images raw,unfilt, fig.width = 10, fig.height = 10}

bmalayi_obj <- LoadImages(doxy_spots, 
                                         time.resolve = F, 
                                         verbose = T)

``` 

## Get image pixels
```{r get image pixels}

# Now we need the pixel coordinates to make it possible to draw points on the capture area
# Most image related info is stored in the "Staffli" object which can be accessed using the
# GetStaffli function

bmalayi_obj.object <- GetStaffli(bmalayi_obj)
bmalayi_obj.object

bmalayi_obj.object@dims[[1]]$width
bmalayi_obj.object@dims[[1]]$height

```

## Set Resolution

```{r set resolution}

# Now we can load the original HE image in whatever resolution we like
# Here: width of 5000 pixels with a fixed aspect ratio
# et a size that is smaller than the original image.

imwidth <- 5000
section <- c(1:5)

```

## Get Image Info

```{r get image info}

# Let's load the image and retrieve some info about it. It will take some time if the image was
# exported in full resolution from the microscope, do NOT plot this image.

im_list <- lapply(1:length(high_res_image_list), function(i) {image_read(high_res_image_list[[i]])})
im_list

```

## Scale Down Image

```{r scale down image }

# An image of 5000 pixels in width is still pretty big to draw in the RStudio Viewer
# so avoid this. Now we can scale down the original image to our
# predefined imwidth and convert it to raster format.

im_list <- lapply(1:length(im_list), function(i) {image_scale(im_list[[i]], paste0(imwidth)) %>% as.raster()})

lapply(1:length(im_list), function(i) {
  print(class(im_list[[i]]))
  print(dim(im_list[[i]]))})

```

## Scale Pixel Coordinates

```{r scale pixel coordinates}

# Next, we need to scale our pixel coordinates to fit our new image. The pixel coordinates
# are stored in the st.object meta.data slot. If multiple samples present here,
# first subset this meta.data to include the section of interest.

gg_list <- lapply(1:length(im_list), function(i) {subset(bmalayi_obj.object@meta.data, sample %in% paste0(section[i]))[, c("pixel_x", "pixel_y")]}) 
gg_list

```

## Get Scaling Factor

```{r get scaling factor}

# We can get the scaling factor by dividing the width of our new image by the
# width of our images that were used when running InputFromTable, i.e. the
# "tissue_hires_image.png".

bmalayi_obj.object@dims[[1]]$width

scale.factor_list <- lapply(1:length(im_list), function(i) {dim(im_list[[i]])[2]/bmalayi_obj.object@dims[[1]]$width})

scale.factor_list

```

## Get New Coordinates

```{r get new coordinates}

# Then we multiply our pixel coordinates by this scale.factor to get our new coordinates
 
gg_list <- Map('*',gg_list,scale.factor_list)
gg_list 

```


# colocal expression images

## binary image

### Add Feature Values

```{r add features values}

# The next thing you need to do is to add the feature values you want to draw.
# We can use the FetchData function which collects
# values from different slots

ftrs <- c("wolbachia")

data_list <- lapply(1:length(gg_list), function(i) {FetchData(object = bmalayi_obj, vars = ftrs, cells = rownames(subset(bmalayi_obj.object@meta.data, sample %in% paste0(section[i]))))})

data_list

gg_list <- Map(cbind , gg_list, data_list)

gg_list 

```

```{r ggplot wol binary}

# Figure 6F

selected.feature <- "wolbachia"
g <- lapply(1:length(section), function(i) grid::rasterGrob(im_list[[i]], width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE))

for(i in seq_along(gg_list)) {
  g_sub = g[[i]]
  
  p <- ggplot() +
  annotation_custom(g_sub, -Inf, Inf, -Inf, Inf) +
  geom_point(data = gg_list[[i]], aes_string("pixel_x", "dim(im_list[[i]])[1] - pixel_y", color = as.factor(gg_list[[i]]$wolbachia)), size = 8) +
  scale_x_continuous(limits = c(0, dim(im_list[[i]])[2]), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, dim(im_list[[i]])[1]), expand = c(0, 0)) +
  theme_void() +
  scale_color_manual(values = c("seagreen2","tan1")) +
  theme(legend.position = c(0.89, 0.85), legend.key.height = unit(x = 1.2, units = "cm"),
        legend.key.width = unit(x = 1.2, units = "cm"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20))
    theme_void() 
  
  pdf(file = paste0("Figure_6F_binary_colocal_", unique(bmalayi_obj$sample_section)[i] ,".pdf"), width = 33.336, height = 32.28) #, res = 300
print(p)
dev.off()
}

```

## Genes on tissue

## Add Feature Values

``` {r genes of interest}

#"bmalayi-WBGene00233912" = "Bma-ubq-2"
#"bmalayi-WBGene00220387" = "Bm126"

genes_for_spatial <- c("bmalayi-WBGene00220387","bmalayi-WBGene00233912")

```


```{r add features values}

# The next thing you need to do is to add the feature values you want to draw.
# Let's select a couple of different ones. We can use the FetchData function which collects
# values from different slots, here from the "meta.data", "assays$SCT@data" and "reductions@UMAP@cell.embeddings" slots


ftrs <- genes_for_spatial
ftrs.edited <- gsub("-", ".", ftrs)

data_list <- lapply(1:length(gg_list), function(i) {FetchData(object = bmalayi_obj, vars = ftrs, cells = rownames(subset(bmalayi_obj.object@meta.data, sample %in% paste0(section[i]))))}) 

data_list

gg_list <- Map(cbind , gg_list, data_list)

gg_list 

```

```{r ggplot wol neg}

# Figure 6F

for (feat in ftrs.edited) {
  selected.feature <- feat
  folder.name <- gsub("bmalayi.", "bmalayi-", feat)
  
  g <- lapply(1:length(section), function(i) grid::rasterGrob(im_list[[i]], width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE))
  
  for(i in seq_along(gg_list)) {
    g_sub = g[[i]]
    df = as.data.frame(gg_list[i])
    df_no_wol <- df[(df$wolbachia == "0"),]
    
    p <- ggplot() +
    annotation_custom(g_sub, -Inf, Inf, -Inf, Inf) +
    geom_point(data = df_no_wol, aes_string("pixel_x", "dim(im_list[[i]])[1] - pixel_y", color = selected.feature), size = 8) +
    scale_x_continuous(limits = c(0, dim(im_list[[i]])[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, dim(im_list[[i]])[1]), expand = c(0, 0)) +
    theme_void() +
    scale_color_gradientn(colors = c("navyblue", "cyan", "yellow", "red", "dark red"), limits = c(0, max(abs(df[, selected.feature])))) +
    theme(legend.position = c(0.89, 0.85), legend.key.height = unit(x = 1.2, units = "cm"),
          legend.key.width = unit(x = 1.2, units = "cm"),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20))
      theme_void() 

    pdf(file = paste0("Figure_6F_", folder.name, "_", unique(bmalayi_obj$sample_section)[i],"_", "no_wol_0",".pdf"), width = 33.336, height = 32.28)
  print(p)
  dev.off()
  }
}


```


```{r ggplot wol pos}

# Figure 6F

for (feat in ftrs.edited) {
  selected.feature <- feat
  folder.name <- gsub("bmalayi.", "bmalayi-", feat)
  
  g <- lapply(1:length(section), function(i) grid::rasterGrob(im_list[[i]], width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE))
  
  for(i in seq_along(gg_list)) {
    g_sub = g[[i]]
    df = as.data.frame(gg_list[i])
    df_wol <- df[(df$wolbachia == "1"),]
    
    p <- ggplot() +
    annotation_custom(g_sub, -Inf, Inf, -Inf, Inf) +
    geom_point(data = df_wol, aes_string("pixel_x", "dim(im_list[[i]])[1] - pixel_y", color = selected.feature), size = 8) +
    scale_x_continuous(limits = c(0, dim(im_list[[i]])[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, dim(im_list[[i]])[1]), expand = c(0, 0)) +
    theme_void() +
    scale_color_gradientn(colors = c("navyblue", "cyan", "yellow", "red", "dark red"), limits = c(0, max(abs(df[, selected.feature])))) +
    theme(legend.position = c(0.89, 0.85), legend.key.height = unit(x = 1.2, units = "cm"),
          legend.key.width = unit(x = 1.2, units = "cm"),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 20))
      theme_void() 

    pdf(file = paste0("Figure_6F_", folder.name, "_", unique(bmalayi_obj$sample_section)[i],"_", "wol_1",".pdf"), width = 33.336, height = 32.28) #, res = 300
  print(p)
  dev.off()
  }
}

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

Output is the genes on tissue plots used in Figure 6F.