---
title: "wolbachia_density_correlation_analysis"
author: "Hailey Sounart"
date: "9/1/2023"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r}
library(Mfuzz)
library(openxlsx)
library(topGO)
library(ggplot2)
library(biomaRt)
library(cowplot)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")

``` 

# Read in Wolbachia abundance data frame
```{r}

wol_abund_df <- readRDS(file = "/Users/hailey.sounart/Documents/brugia_malayi/220815_final_code/230214_repeat_with_new_section_names/wol_abund_group_avg_df.rds")

```

# Make ExpressionSet object
```{r}

eset <- new('ExpressionSet', exprs=(as.matrix(wol_abund_df)))

```

# Check for missing values
```{r}

eset.r <- filter.NA(eset, thres=0.25)
eset.f <-fill.NA(eset.r,mode="knnw")

```

# Filtering
```{r}

#This function is to exclude genes with low standard deviation.
#min.std=0 means filtering out genes that have no variation across the Wolbachia abundance groups, i.e., genes that have a constant expression profile. A standard deviation of 0 indicates that all the expression values for that gene are the same across all Wolbachia abundance groups.

eset.filt <- filter.std(eset.f,min.std=0)

```

# Standardization
```{r}

eset.s <- standardise(eset.filt)

```

# Test for m (fuzzifier) value
```{r}

m.s <- mestimate(eset.s)
m.s

```

# Check for cluster numbers

```{r}

Dmin(eset.s,m.s ,crange=seq(3,20,1),repeats=3,visu=TRUE)

```
Dmin declines slower after 6, will try 3-6 clusters.

# Soft clustering
```{r}
set.seed(45)

clu_4 <- mfuzz(eset.s, c=4, m=m.s)

```

# View clusters
```{r}
set.seed(45)

group_names <- colnames(exprs(eset.s))
mfuzz.plot(eset.s,cl=clu_4,time.labels=group_names, new.window=FALSE, mfrow=c(2,2))
```
# Core cluster genes
```{r}

clu_genes <- acore(eset.s,cl=clu_4,min.acore=0.7)
clu_genes

```

# BioMart

```{r}

bmalayi_exp_norm_genes <- read.xlsx("bmalayi_exp_norm_genes.xlsx", colNames = FALSE)
bmalayi_exp_norm_genes_list <- bmalayi_exp_norm_genes$X1
#length(bmalayi_exp_norm_genes_list)
#head(bmalayi_exp_norm_genes_list)

```


```{r}
mart <- useMart("parasite_mart", dataset = "wbps_gene", host = "https://parasite.wormbase.org", port = 443)

```

```{r}
listDatasets(mart)
```

```{r}

all_gene_GO <- getBM(mart = mart, 
                     filters = c("species_id_1010", "wbps_gene_id"),
                     value = list("brmalaprjna10729", bmalayi_exp_norm_genes_list),
                     attributes = c("production_name_1010", "wbps_gene_id", "display_name_1010", "species_id_key", "taxonomy_id_1010", "assembly_accession_1010", "nematode_clade_1010" ,"wbps_transcript_id","wbps_peptide_id", "external_gene_id", "external_gene_db", "description", "gene_biotype", "external_transcript_id", "transcript_db_name", "transcript_biotype", "entrezgene_id", "entrezgene_name", "wormbase_gene", "wormbase_transcript", "go_accession", "go_name_1006", "go_definition_1006", "go_linkage_type", "go_namespace_1003"))

```

# GO term enrichment analysis

## Get all experiment genes (background genes)

```{r}

#all exp genes

# Create a list where each gene is associated with its GO terms
gene2GO <- split(all_gene_GO$go_accession, all_gene_GO$wbps_gene_id)

gene2GO <- gene2GO[sapply(gene2GO, length) > 0]

# All genes in the background (assuming all genes in df are the background)
allGenes <- unique(all_gene_GO$wbps_gene_id)
```

## cluster genes
```{r}

# cluster 1
selectedGenes_clu_1 <- rownames(clu_genes[[1]])
selectedGenes_clu_1 <- gsub("bmalayi-", "",selectedGenes_clu_1)

# cluster 2
selectedGenes_clu_2 <- rownames(clu_genes[[2]])
selectedGenes_clu_2 <- gsub("bmalayi-", "",selectedGenes_clu_2)

# cluster 3
selectedGenes_clu_3 <- rownames(clu_genes[[3]])
selectedGenes_clu_3 <- gsub("bmalayi-", "",selectedGenes_clu_3)

# cluster 4
selectedGenes_clu_4 <- rownames(clu_genes[[4]])
selectedGenes_clu_4 <- gsub("bmalayi-", "",selectedGenes_clu_4)

```

```{r}

# Create a binary gene list indicating if each gene is in the cluster of interest
#cluster 1
geneList_clu_1 <- factor(as.integer(allGenes %in% selectedGenes_clu_1), levels = c(0, 1))
names(geneList_clu_1) <- allGenes

#cluster 2
geneList_clu_2 <- factor(as.integer(allGenes %in% selectedGenes_clu_2), levels = c(0, 1))
names(geneList_clu_2) <- allGenes

# cluster 3
geneList_clu_3 <- factor(as.integer(allGenes %in% selectedGenes_clu_3), levels = c(0, 1))
names(geneList_clu_3) <- allGenes

# cluster 4
geneList_clu_4 <- factor(as.integer(allGenes %in% selectedGenes_clu_4), levels = c(0, 1))
names(geneList_clu_4) <- allGenes

```

## run topGO

```{r}
#cluster 1
clu_1_GOdata_BP <- new("topGOdata", description="Pattern_1", ontology="BP", allGenes = geneList_clu_1, annot = annFUN.gene2GO, gene2GO = gene2GO) 

clu_1_resultFisher_BP <- runTest(clu_1_GOdata_BP, algorithm="classic", statistic="fisher") 

results_clu_1_BP <- GenTable(clu_1_GOdata_BP, classicFisher = clu_1_resultFisher_BP, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 20, numChar = 5000)

# Extract p-values from the resultFisher object
resultScores_clu1 <- score(clu_1_resultFisher_BP)
      
# Add the p-values as a new column to the results dataframe
results_clu_1_BP$p_value <- resultScores_clu1[results_clu_1_BP$GO.ID]
results_clu_1_BP
```

```{r}
#cluster 2
clu_2_GOdata_BP <- new("topGOdata", description="Pattern_2", ontology="BP", allGenes = geneList_clu_2, annot = annFUN.gene2GO, gene2GO = gene2GO) 

clu_2_resultFisher_BP <- runTest(clu_2_GOdata_BP, algorithm="classic", statistic="fisher") 

results_clu_2_BP <- GenTable(clu_2_GOdata_BP, classicFisher = clu_2_resultFisher_BP, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 20, numChar = 5000)

# Extract p-values from the resultFisher object
resultScores_clu2 <- score(clu_2_resultFisher_BP)
      
# Add the p-values as a new column to the results dataframe
results_clu_2_BP$p_value <- resultScores_clu2[results_clu_2_BP$GO.ID]
results_clu_2_BP
```

```{r}
#cluster 3
clu_3_GOdata_BP <- new("topGOdata", description="Pattern_3", ontology="BP", allGenes = geneList_clu_3, annot = annFUN.gene2GO, gene2GO = gene2GO) 

clu_3_resultFisher_BP <- runTest(clu_3_GOdata_BP, algorithm="classic", statistic="fisher") 

results_clu_3_BP <- GenTable(clu_3_GOdata_BP, classicFisher = clu_3_resultFisher_BP, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 20, numChar = 5000)

# Extract p-values from the resultFisher object
resultScores_clu3 <- score(clu_3_resultFisher_BP)
      
# Add the p-values as a new column to the results dataframe
results_clu_3_BP$p_value <- resultScores_clu3[results_clu_3_BP$GO.ID]
results_clu_3_BP
```

```{r}
#cluster 4
clu_4_GOdata_BP <- new("topGOdata", description="Pattern_4", ontology="BP", allGenes = geneList_clu_4, annot = annFUN.gene2GO, gene2GO = gene2GO) 

clu_4_resultFisher_BP <- runTest(clu_4_GOdata_BP, algorithm="classic", statistic="fisher") 

results_clu_4_BP <- GenTable(clu_4_GOdata_BP, classicFisher = clu_4_resultFisher_BP, orderBy = "resultFisher", ranksOf = "classicFisher", topNodes = 20, numChar = 5000)

# Extract p-values from the resultFisher object
resultScores_clu4 <- score(clu_4_resultFisher_BP)
      
# Add the p-values as a new column to the results dataframe
results_clu_4_BP$p_value <- resultScores_clu4[results_clu_4_BP$GO.ID]
results_clu_4_BP
```

## view GO results

```{r}
GO_list <- list(results_clu_1_BP, results_clu_2_BP, results_clu_3_BP, results_clu_4_BP)
```


```{r}

#GO enrichment plots for each cluster

plot_list <- lapply(GO_list, function(cluster_results) {
  
  # Create a data frame with GO term and -log10(p-value)
  plot_data <- data.frame(GO.Term = paste(cluster_results$GO.ID, cluster_results$Term, sep = ", "),
                          neg_log_pval = -log10(cluster_results$p_value))
  
  # Order the data frame by -log10(p-value)
  plot_data <- plot_data[order(plot_data$neg_log_pval, decreasing = TRUE), ]
  top_10 <- head(plot_data, 10)
  
  # Create a bar plot using ggplot2
  plot <- ggplot(top_10, aes(x = neg_log_pval, y = reorder(GO.Term, neg_log_pval))) +
    geom_bar(stat = "identity", aes(fill = "#EF5B5B")) +
    labs(x = "-log10(p-value)", y = "GO Term") +
    geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "black") +
    theme_minimal() +
    theme(axis.text.y = element_text(hjust = 1, color = "black"),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.ticks = element_line(color = "black"),
          axis.line = element_line(color = "black"),
          plot.title = element_text(hjust = 0.5),
          panel.background = element_blank(),  # Remove panel background
          axis.ticks.length = unit(0.25, "cm"),  # Adjust the length of the tick marks if needed
          axis.text.x = element_text(color = "black")) +  # Make x axis text black
    coord_cartesian(xlim = c(0, 2.5))
    
  return(plot)
})

plot_list

```

# Save results

## MFuzz
```{r}

#Supplementary Data 6

clu_gene_core_list <- lapply(clu_genes, function(clu) {
  
  clu$NAME <- gsub("bmalayi-", "",clu$NAME)
  
  clu <- clu[, c("NAME", "MEM.SHIP")]
  
  colnames(clu) <- c("Gene", "Membership")
  
  return(clu)
})

clu_gene_core_list

# Specify the Excel file name
excel_file <- "Supplementary_Data_6.xlsx"

# Create a new workbook
wb_clu <- createWorkbook()

# Add data frames as sheets
addWorksheet(wb_clu, "Pattern_1")
writeData(wb_clu, "Pattern_1", clu_gene_core_list[[1]])

addWorksheet(wb_clu, "Pattern_2")
writeData(wb_clu, "Pattern_2", clu_gene_core_list[[2]])

addWorksheet(wb_clu, "Pattern_3")
writeData(wb_clu, "Pattern_3", clu_gene_core_list[[3]])

addWorksheet(wb_clu, "Pattern_4")
writeData(wb_clu, "Pattern_4", clu_gene_core_list[[4]])

# Save workbook to file
saveWorkbook(wb_clu, "Supplementary_Data_6.xlsx", overwrite = TRUE)
```

```{r}
#Supplementary Figure 2D

group_names <- colnames(exprs(eset.s))

pdf("Supplementary_Figure_2D.pdf", width = 10, height = 7, pointsize = 12)

mfuzz.plot2(eset.s, cl=clu_4, time.labels=group_names, x11=FALSE, xlab="Wolbachia abundance group",mfrow=c(2,2), colo = "fancy")

dev.off()
```

```{r}

#Supplementary Figure 2D

pdf("Supplementary_Figure_2D_bar.pdf", width = 10, height = 10, pointsize = 12)

mfuzzColorBar(col="fancy",main="Membership",cex.main=1, horizontal=TRUE)

dev.off()
```

## topGO 

```{r}
#save topGo plots
arranged_plots <- plot_grid(plotlist = plot_list, ncol = 2, nrow = 2, align = "v")
ggsave(file = "Supplementary_Figure_2E.pdf", plot = arranged_plots, dpi = 300, width = 17, height = 5) #Supplementary Figure 2E

```

```{r}

#Supplementary Data 7

#GO terms table for each cluster (top 30 terms)

# Specify the Excel file name
excel_file <- "Supplementary_Data_7.xlsx"

# Create a new workbook
wb <- createWorkbook()

# Add data frames as sheets
addWorksheet(wb, "Pattern_1")
writeData(wb, "Pattern_1", GO_list[[1]])

addWorksheet(wb, "Pattern_2")
writeData(wb, "Pattern_2", GO_list[[2]])

addWorksheet(wb, "Pattern_3")
writeData(wb, "Pattern_3", GO_list[[3]])

addWorksheet(wb, "Pattern_4")
writeData(wb, "Pattern_4", GO_list[[4]])

# Save workbook to file
saveWorkbook(wb, "Supplementary_Data_7.xlsx", overwrite = TRUE)


```

# Check for genes

## glycolysis genes
```{r glycolysis genes names}

glycolysis_genes <- c("WBGene00234007","WBGene00226800","WBGene00234111","WBGene00227349","WBGene00227350","WBGene00228818","WBGene00223396","WBGene00225841","WBGene00234141","WBGene00225232","WBGene00225960","WBGene00234100","WBGene00233578","WBGene00234226","WBGene00225601","WBGene00226459","WBGene00223851","WBGene00228293","WBGene00230482","WBGene00230483","WBGene00223624","WBGene00224939","WBGene00230488","WBGene00233072","WBGene00269278","WBGene00223126","WBGene00228134","WBGene00226573","WBGene00224695","WBGene00225076","WBGene00225490","WBGene00226192","WBGene00226849","WBGene00229624","WBGene00223600","WBGene00226903","WBGene00229687","WBGene00233045","WBGene00231340","WBGene00233943","WBGene00228214","WBGene00229261","WBGene00223797","WBGene00234093","WBGene00227890","WBGene00225448","WBGene00223299","WBGene00225708")
  
glycolysis_genes_rename <- paste0("bmalayi-", glycolysis_genes)

```

```{r}

#glycolytic genes found in each cluster

gly_overlap_list <- lapply(clu_gene_core_list, function(cluster_results) {
  
  inter <- intersect(rownames(cluster_results), glycolysis_genes_rename)
  
  return(cluster_results[inter,])
    
})

gly_overlap_list

```

## Colocal genes
```{r}

coloca_de_genes <- readRDS(file = "/Users/hailey.sounart/Documents/brugia_malayi/220815_final_code/230214_repeat_with_new_section_names/brugia_colocal_wilcox_deseq2_p0.05_fc0.1_de_genes.rds")

coloca_de_genes_rename <- paste0("bmalayi-", rownames(coloca_de_genes))

```

```{r}

#colocalization genes found in each cluster

colocal_overlap_list <- lapply(clu_gene_core_list, function(cluster_results) {
  
  inter <- intersect(rownames(cluster_results), coloca_de_genes_rename)
  
  return(cluster_results[inter,])
  
    
})

colocal_overlap_list

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

Output is Supplementary Figure 2D-E plots and Supplementary Data 6,7. 
