---
title: "Make Brugia malayi and Wolbachia (wBm) gene lists"
author: "Hailey Sounart"
date: "8/15/2022"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time1 <- Sys.time()
start_time1
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(stringr)
  library(openxlsx)
  library(tidyr)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")

``` 

## Get Brugia malayi genes

```{r read in bmalayi annotation}

#read in the bmalayi annotation file as a table
all_bmalayi_gene_annotations <- read.table("brugia_malayi.gene_biotype_product.map.tsv", sep="\t", header=T, stringsAsFactors = F)
all_bmalayi_gene_annotations

```

```{r get all bmalayi genes}

#get gene ids from annotation table
bmalayi_genes <- all_bmalayi_gene_annotations$gene_id

#add bmalayi- in front so matches syntax of gene names in the count matrices
bmalayi_genes <- paste("bmalayi-", bmalayi_genes, sep="")

```

## Save brugia malayi all genes list

```{r save all genes list}

length(bmalayi_genes)

dir.create(paste0("bmalayi_all_genes"))

saveRDS(bmalayi_genes, file = "bmalayi_all_genes/bmalayi_genes.rds")

write.csv(bmalayi_genes, file = "bmalayi_all_genes/bmalayi_genes.csv")

write.xlsx(bmalayi_genes, file = "bmalayi_all_genes/bmalayi_genes.xlsx", col.names=TRUE, row.names=TRUE)
```


## Get Brugia malayi rRNA genes

Get a list of rRNA genes
```{r get rRNA genes}

#get bmalayi genes annotated as rRNA in the biotype column
bmalayi_rRNA <- (all_bmalayi_gene_annotations$biotype == "rRNA")

#subset bmalayi annotation table for rRNA genes
bmalayi_rRNA_table <- all_bmalayi_gene_annotations[bmalayi_rRNA,]
bmalayi_rRNA_table 

#get rRNA gene ids from rRNA annotation table
bmalayi_rRNA_genes <- bmalayi_rRNA_table$gene_id

#add bmalayi- in front so matches syntax of gene names in the count matrices
bmalayi_rRNA_genes <- paste("bmalayi-", bmalayi_rRNA_genes, sep="")

```

## Get ribosomal genes

```{r see ribosomal protein coding genes}

#see product column values that contain ribosomal protein coding annotations in the bmalayi annotation table

ribo_string_variations <- c(".*ribosomal.*", ".*Ribosomal.*", ".*ribosome.*", ".*Ribosome.*")

ribo_name_lists <- lapply(1:length(ribo_string_variations), function(i) {grep(pattern = ribo_string_variations[[i]], x = all_bmalayi_gene_annotations$product, value = TRUE)})

```

```{r get ribosomal protein coding genes 1}

#get bmalayi genes annotated as ribosomal in the biotype column

bmalayi_ribo_list <- lapply(1:length(ribo_string_variations), function(i) {str_detect(all_bmalayi_gene_annotations$product, regex(ribo_string_variations[[i]], dotall = TRUE))})

bmalayi_ribo_list_no_na <- lapply(1:length(bmalayi_ribo_list), function(i) {bmalayi_ribo_list[[i]] %>% replace_na(FALSE)})
```

```{r get ribosomal protein coding genes 2}

#subset annotation table for ribosomal/Ribosomal genes

ribo_table_list <- lapply(1:length(bmalayi_ribo_list_no_na), function(i) {all_bmalayi_gene_annotations[bmalayi_ribo_list_no_na[[i]],]})

ribo_table_list

```

```{r get ribosomal protein coding genes 3}

#get ribo gene lists and add bmalayi- in front so matches syntax of gene names in the count matrices

ribo_genes_list <- lapply(1:length(ribo_table_list), function(i) {ribo_table_list[[i]]$gene_id})
ribo_genes_list <- lapply(1:length(ribo_genes_list), function(i) {paste("bmalayi-", ribo_genes_list[[i]], sep="")})

```

```{r get ribosomal protein coding genes 4}

#combine ribosomal and Ribosomal gene lists

all_ribo_genes <- unlist(ribo_genes_list, recursive = TRUE, use.names = TRUE)

```

## Save rRNA and ribosomal gene lists

```{r save rRNA genes}

#bmalayi_rRNA_genes

dir.create(paste0("rRNA_genes"))

saveRDS(bmalayi_rRNA_genes, file = "rRNA_genes/bmalayi_rRNA_genes.rds")

write.csv(bmalayi_rRNA_genes, file = "rRNA_genes/bmalayi_rRNA_genes.csv")

write.xlsx(bmalayi_rRNA_genes, file = "rRNA_genes/bmalayi_rRNA_genes.xlsx", col.names=TRUE, row.names=TRUE)
```

```{r save ribosomal genes}

#all_ribo_genes

dir.create(paste0("ribosomal_genes"))

saveRDS(all_ribo_genes, file = "ribosomal_genes/bmalayi_ribo_genes.rds")

write.csv(all_ribo_genes, file = "ribosomal_genes/bmalayi_ribo_genes.csv")

write.xlsx(all_ribo_genes, file = "ribosomal_genes/bmalayi_ribo_genes.xlsx", col.names=TRUE, row.names=TRUE)

```

## Combine rRNA and ribosomal genes into a single list
```{r save rRNA and ribosomal genes}

bmalayi_rRNA_ribo_genes <- c(bmalayi_rRNA_genes, all_ribo_genes)
#bmalayi_rRNA_ribo_genes

dir.create(paste0("rRNA_and_ribo_genes"))

saveRDS(bmalayi_rRNA_ribo_genes, file = "rRNA_and_ribo_genes/bmalayi_rRNA_and_ribo_genes.rds")

write.csv(bmalayi_rRNA_ribo_genes, file = "rRNA_and_ribo_genes/bmalayi_rRNA_and_ribo_genes.csv")

write.xlsx(bmalayi_rRNA_ribo_genes, file = "rRNA_and_ribo_genes/bmalayi_rRNA_and_ribo_genes.xlsx", col.names=TRUE, row.names=TRUE)

```

## Get Brugia malayi mitochondiral genes

```{r see mitochondrial protein coding genes}

#see product column values that contain mitochondrial or Mitochondrial in the bmalayi annotation table

mito_string_variations <- c(".*mitoc.*", ".*Mitoc.*")

mito_name_lists <- lapply(1:length(mito_string_variations ), function(i) {grep(pattern =mito_string_variations[[i]], x = all_bmalayi_gene_annotations$product, value = TRUE)})

```

```{r get mitochondrial protein coding genes 1}

#get bmalayi genes annotated as mitochondrial or Mitochondrial in the biotype column

bmalayi_mito_list <- lapply(1:length(mito_string_variations), function(i) {str_detect(all_bmalayi_gene_annotations$product, regex(mito_string_variations[[i]], dotall = TRUE))})

bmalayi_mito_list_no_na <- lapply(1:length(bmalayi_mito_list), function(i) {bmalayi_mito_list[[i]] %>% replace_na(FALSE)})

```

```{r get mitochondrial protein coding genes 2}

#subset annotation table for mitochondrial/Mitochondrial genes

mito_table_list <- lapply(1:length(bmalayi_mito_list_no_na), function(i) {all_bmalayi_gene_annotations[bmalayi_mito_list_no_na[[i]],]})

mito_table_list

```

```{r get mitochondrial protein coding genes 3}

#get mito gene lists and add bmalayi- in front so matches syntax of gene names in the count matrices

mito_genes_list <- lapply(1:length(mito_table_list), function(i) {mito_table_list[[i]]$gene_id})
mito_genes_list <- lapply(1:length(mito_genes_list), function(i) {paste("bmalayi-", mito_genes_list[[i]], sep="")})

```

```{r get mitochondrial protein coding genes 4}

#combine mitochondrial and Mitochondrial gene lists

all_mito_genes <- unlist(mito_genes_list, recursive = TRUE, use.names = TRUE)

```

# Save mitochondrial lists

```{r save mito genes}

#all_mito_genes

dir.create(paste0("mitochondrial_genes"))

saveRDS(all_mito_genes, file = "mitochondrial_genes/bmalayi_mito_genes.rds")

write.csv(all_mito_genes, file = "mitochondrial_genes/bmalayi_mito_genes.csv")

write.xlsx(all_mito_genes, file = "mitochondrial_genes/bmalayi_mito_genes.xlsx", col.names=TRUE, row.names=TRUE)

```

## Get Brugia malayi protein coding genes

```{r get bmalayi protein coding genes}

# Get Brugia malayi protein coding gene list
bmalayi_coding <- (all_bmalayi_gene_annotations$biotype == "protein_coding")

bmalayi_coding_list <- str_detect(all_bmalayi_gene_annotations$biotype, regex(".*protein_coding.*", dotall = TRUE))
bmalayi_coding_list[is.na(bmalayi_coding_list)] <- FALSE

bmalayi_coding_table <- all_bmalayi_gene_annotations[bmalayi_coding_list,]

bmalayi_coding_genes <- bmalayi_coding_table$gene_id

bmalayi_coding_genes <- paste("bmalayi-", bmalayi_coding_genes, sep="")
length(bmalayi_coding_genes)

```

# Save Brugia malayi protein coding gene list

```{r save bmalayi protein coding genes}

dir.create(paste0("all_protein_coding_genes"))

saveRDS(bmalayi_coding_genes, file = "all_protein_coding_genes/bmalayi_protein_coding_genes.rds")

write.csv(bmalayi_coding_genes, file = "all_protein_coding_genes/bmalayi_protein_coding_genes.csv")

write.xlsx(bmalayi_coding_genes, file = "all_protein_coding_genes/bmalayi_protein_coding_genes.xlsx", col.names=TRUE, row.names=TRUE)

```

## bmalayi protein coding (no ribo genes)

```{r remove ribosomal genes from bmalayi protein coding genes}

#remove ribosomal genes from bmalayi protein coding genes
bmalayi_protein_coding.no.ribo <- bmalayi_coding_genes[!(bmalayi_coding_genes %in% all_ribo_genes)]
length(bmalayi_protein_coding.no.ribo)


#check removed ribosomal genes
intersect(bmalayi_protein_coding.no.ribo,all_ribo_genes)

``` 

## Save bmalayi protein coding (no ribo)

```{r save bmalayi protein coding (no ribo)}

dir.create(paste0("bmalayi_protein_coding_no_ribo"))

saveRDS(bmalayi_protein_coding.no.ribo, file = "bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.rds")

write.csv(bmalayi_protein_coding.no.ribo, file = "bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.csv")

write.xlsx(bmalayi_protein_coding.no.ribo, file = "bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.xlsx", col.names=TRUE, row.names=TRUE)

```

## bmalayi protein coding (no mito genes)

```{r remove mitochondrial genes from bmalayi protein coding genes}

#remove mitochondrial genes from bmalayi protein coding genes
bmalayi_protein_coding.no.mito <- bmalayi_coding_genes[!(bmalayi_coding_genes %in% all_mito_genes)]
length(bmalayi_protein_coding.no.mito)

#check removed mitochondrial genes
intersect(bmalayi_protein_coding.no.mito,all_mito_genes)

``` 

# Save bmalayi protein coding (no mito)

```{r save bmalayi protein coding (no mito)}

dir.create(paste0("bmalayi_protein_coding_no_mito"))

saveRDS(bmalayi_protein_coding.no.mito, file = "bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.rds")

write.csv(bmalayi_protein_coding.no.mito, file = "bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.csv")

write.xlsx(bmalayi_protein_coding.no.mito, file = "bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.xlsx", col.names=TRUE, row.names=TRUE)

```

## bmalayi protein coding (no ribo+mito genes)

```{r remove ribosomal and mitochondrial genes from bmalayi protein coding genes}

length(intersect(all_ribo_genes,all_mito_genes))
all_ribo_mito_genes <- c(all_ribo_genes,all_mito_genes)
all_ribo_mito_genes <- unique(all_ribo_mito_genes)
length(all_ribo_mito_genes)

length(bmalayi_coding_genes)
length(intersect(bmalayi_coding_genes,all_mito_genes))
length(intersect(bmalayi_coding_genes,all_ribo_genes))

#remove ribosomal, mitochondrial genes from bmalayi protein coding genes
bmalayi_protein_coding.no.ribo.no.mito <- bmalayi_coding_genes[!(bmalayi_coding_genes %in% all_ribo_mito_genes)]
length(bmalayi_protein_coding.no.ribo.no.mito)


#check removed ribosomal, mitochondrial genes
intersect(bmalayi_protein_coding.no.ribo.no.mito,all_ribo_mito_genes)

``` 

# Save bmalayi protein coding (no ribo+mito genes)

```{r save bmalayi protein coding (no ribo+mito)}

dir.create(paste0("bmalayi_protein_coding_no_ribo_no_mito"))

saveRDS(bmalayi_protein_coding.no.ribo.no.mito, file = "bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")

write.csv(bmalayi_protein_coding.no.ribo.no.mito, file = "bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.csv")

write.xlsx(bmalayi_protein_coding.no.ribo.no.mito, file = "bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.xlsx", col.names=TRUE, row.names=TRUE)

```

## Get Wolbachia (wBm) genes

```{r get wBm genes}

#For Wolbachia
all_wBm_gene_annotations <- read.delim("wBm.gene_biotype_product.map.tsv", sep="\t", header=F, stringsAsFactors = F)
all_wBm_gene_annotations

wBm_coding_list <- str_detect(all_wBm_gene_annotations$V2, regex(".*protein_coding.*", dotall = TRUE))
wBm_coding_list[is.na(wBm_coding_list)] <- FALSE
#wBm_coding_list

wBm_coding_table <- all_wBm_gene_annotations[wBm_coding_list,]
wBm_coding_table

wBm_coding_genes <- wBm_coding_table$V1
#wBm_coding_genes

wBm_coding_genes <- paste("wBm-----", wBm_coding_genes, sep="")
wBm_coding_genes <- gsub("_", "-", wBm_coding_genes)
#head(wBm_coding_genes)
length(wBm_coding_genes)

all_wBm_genes <- all_wBm_gene_annotations$V1
all_wBm_genes <- paste("wBm-----", all_wBm_genes, sep="")
all_wBm_genes <- gsub("_", "-", all_wBm_genes)
length(all_wBm_genes)
#head(all_wBm_genes)

```

# Save wBm gene lists

```{r save wBm genes}

dir.create(paste0("wBm_all_genes"))

saveRDS(all_wBm_genes, file = "wBm_all_genes/all_wBm_genes.rds")

write.csv(all_wBm_genes, file = "wBm_all_genes/all_wBm_genes.csv")

write.xlsx(all_wBm_genes, file = "wBm_all_genes/all_wBm_genes.xlsx", col.names=TRUE, row.names=TRUE)

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time1 <- Sys.time()
end_time1
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time1 <- round((end_time1-start_time1), 3)
elapsed_time1
```

Outputs are all the genes lists saved in 3 formats: rds object, excel file, and csv file. The output gene lists are of all brugia genes, brugia rRNA and ribosomal genes, brugia mitochondrial genes, brugia protein coding genes, brugia protein coding genes without ribosomal genes, brugia protein coding genes without mitochondrial genes, brugia protein coding genes without mitochondrial and ribosomal genes, and Wolbachia (wBm) genes. 
