---
title: "Load counts matrices and add metadata"
author: "Hailey Sounart"
date: "8/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

## Load packages
```{r load packages}
  library(STutility)
  library(hdf5r)
  library(imager)
  library(akima)
  library(plotly)
  library(raster)
  library(dplyr)
  library(magrittr)
  library(ggplot2)
  library(imager) 
  library(magrittr)
  library(RColorBrewer)
  library(stringr)
  library(openxlsx)
  library(hash)
  library(readxl)
  library(purrr)
  library(harmony)
```

```{r set working directory}

setwd("<path to the directory where count matrices for all samples reside>")
``` 

## Load processed data (spaceranger outputs)
```{r Load filtered data}

all_sections_subset_infotable_for_labels <- read.table("infotable.csv", sep=",", header=T, stringsAsFactors = F)
all_sections_subset_infotable_for_labels

``` 

## Make Seurat object
```{r make object}

all_sections_subset_for_labels <- InputFromTable(infotable = all_sections_subset_infotable_for_labels,
                                 min.gene.count = 0, 
                                 min.gene.spots = 0,
                                 min.spot.count = 0,
                                 min.spot.feature.count = 0,
                                 platform="Visium")
``` 

## Section number - Manual Annotations

Manually label each worm section with the section number. See https://giacomellolabst.shinyapps.io/brugiast-shiny/ for the worm sections with the corresponding section number. 

```{r label each worm sample and section, fig.width = 10, fig.height = 10, out.height = "100%", out.width = "100%"}

all_sections_subset_for_labels <- ManualAnnotation(all_sections_subset_for_labels)

``` 

Transfer label to new metadata column called "section".

```{r transfer label for each worm section, fig.width = 10, fig.height = 10, out.height = "100%", out.width = "100%"}

all_sections_subset_for_labels$section <- paste(all_sections_subset_for_labels$labels)
all_sections_subset_for_labels@meta.data

``` 

## Remove any unlabeled soots
```{r remove unlabeled spots, fig.width = 10, fig.height = 10, out.height = "100%", out.width = "100%"}

all_sections_subset_labeled.on_labeled_spot <- SubsetSTData(all_sections_subset_for_labels, expression = (section != "Default"))
all_sections_subset_labeled.on_labeled_spot@meta.data
``` 

## Save Labeled Object

``` {r save labeled object}

saveRDS(all_sections_subset_labeled.on_labeled_spot, file = "bmalayi.manually.section.labeled.rds")

```

See "V10F24_041_spots_seurat_metadata.xlsx" in the github directory "https://github.com/giacomellolab/Brugia_malayi_study/tree/main/data" for how the spot section metadata ends up

## Read in Labeled Object

``` {r save labeled object}

all_sections_subset.section.labeled <- readRDS("bmalayi.manually.section.labeled.rds")

```

Look at object
``` {r save labeled object}

all_sections_subset.section.labeled

```

## Add additional metadata

Add: sample, sub-array, slide, worm batch, ROI, embedding technique, treatment condition, combined label slide_subarray, combined label sample_section, combined label slide_subarray_sample_section

## Sampleid 

Create sampleid number metadata column
```{r sampleid metdata}

#also create sample_id metadata
all_sections_subset.section.labeled$sample_id <- paste0("", 1:9)[as.numeric(do.call(rbind, strsplit(colnames(all_sections_subset.section.labeled), split = "_"))[, 2])]

all_sections_subset.section.labeled@meta.data

```

## Sub-array

Create sub-array metdadata column
``` {r sub-array labels}

all_sections_subset.section.labeled$sub_array <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("A1","B1","D1")
)

all_sections_subset.section.labeled@meta.data

```

## Slide

Create slide metdadata column
``` {r slide labels}

all_sections_subset.section.labeled$slide <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("V10F24_041","V10F24_041","V10F24_041")
)

all_sections_subset.section.labeled@meta.data

```

## Worm batch

Create worm batch metadata column
``` {r worm batch labels}

all_sections_subset.section.labeled$sample_batch <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("3","3","3")
)

all_sections_subset.section.labeled@meta.data

```

## ROI

Create worm ROI metadata column
``` {r ROI labels}

all_sections_subset.section.labeled$ROI <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("1","1","1")
)

all_sections_subset.section.labeled@meta.data

```

## Embedding technique

Create embedding technique metadata column
``` {r emebedding technique labels}

all_sections_subset.section.labeled$embed <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("re_embed","re_embed","original")
)

all_sections_subset.section.labeled@meta.data

```
## Worm sample

Create worm sample metadata column
``` {r worm sample labels}

all_sections_subset.section.labeled$worm_sample <- plyr::mapvalues(
    x = all_sections_subset.section.labeled$sample_id, 
    from = c("1","2","3"),
    to = c("BM1","BM2","BM3")
)

all_sections_subset.section.labeled@meta.data

```

## Combined slide_subarray

```{r echo=FALSE}

all_sections_subset.section.labeled$slide_subarray <- paste(all_sections_subset.section.labeled$slide, all_sections_subset.section.labeled$sub_array, sep="_")

```

## Combined sample_section 

```{r echo=FALSE}

all_sections_subset.section.labeled$sample_section <- paste(all_sections_subset.section.labeled$worm_sample, all_sections_subset.section.labeled$section, sep="_")

```

## Combined slide_subarray_sample_section

```{r echo=FALSE}

all_sections_subset.section.labeled$slide_subarray_sample_section <- paste(all_sections_subset.section.labeled$slide, all_sections_subset.section.labeled$sub_array, all_sections_subset.section.labeled$worm_sample, all_sections_subset.section.labeled$section, sep="_")

```

``` {r check metadata}

all_sections_subset.section.labeled@meta.data

```

## Save Object

``` {r save labeled object}

saveRDS(all_sections_subset.section.labeled, file = "bmalayi.all.labeles.rds")

```

## Save metadata

``` {r save labeled object with metadata}

write.csv(all_sections_subset.section.labeled@meta.data[,], file = "V10F24_041_metadata.csv")

write.xlsx(all_sections_subset.section.labeled@meta.data[,], file = "V10F24_041_metadata.xlsx", col.names=TRUE, row.names=TRUE)

```

## Read in Object

``` {r read in labeled object}

all_sections_subset_QC <- readRDS("bmalayi.all.labeles.rds")

```

## Ribosomal content

```{r load in ribosomal genes, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#load in the list(s) of ribosomal genes

#load in rRNA genes
bmalayi_rRNA_gene_list <- readRDS("rRNA_genes/bmalayi_rRNA_genes.rds") #refhook = NULL

#load in ribosomal protein coding genes
bmalayi_ribo_gene_list <- readRDS("ribosomal_genes/bmalayi_ribo_genes.rds") #refhook = NULL

#combine rRNA and ribosomal protein coding genes into a single list
bmalayi_all_ribo_list_210622 <- c(bmalayi_rRNA_gene_list, bmalayi_ribo_gene_list)

```

```{r get count and percent ribosomal gene expression, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#get the list of genes in the object that intersect/match the genes in the combined ribo list
ribo_in_raw_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_all_ribo_list_210622))
#ribo_in_raw_obj_210518
length(ribo_in_raw_obj_210622)

#create a column "percent.ribo" in the metadata that contains the percentage of ribosomal gene UMI counts in each spot
all_sections_subset_QC$percent.ribo <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[ribo_in_raw_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.ribo" in the metadata that contains the UMI counts of ribosomal genes in each spot
all_sections_subset_QC$counts.ribo <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[ribo_in_raw_obj_210622, ])

```

# rRNA content 

```{r get count and percent rRNA gene expression, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#get the list of genes in the object that intersect/match the genes in the rRNA list
rRNA_in_raw_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_rRNA_gene_list))
#ribo_in_raw_obj_210518
length(rRNA_in_raw_obj_210622)

#create a column "percent.rRNA" in the metadata that contains the percentage of rRNA gene UMI counts in each spot
all_sections_subset_QC$percent.rRNA <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[rRNA_in_raw_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.rRNA" in the metadata that contains the UMI counts of rRNA genes in each spot
all_sections_subset_QC$counts.rRNA <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[rRNA_in_raw_obj_210622, ])

```

## Mitochondrial content

```{r load in mitochondrial genes, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#load in mitochondrial gene list
bmalayi_mito_gene_list <- readRDS("mitochondrial_genes/bmalayi_mito_genes.rds") #refhook = NULL

```

```{r get count and percent mitochondrial gene expression, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#get the list of genes in the object that intersect/match the genes in the mito list
mito_in_raw_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_mito_gene_list))
#ribo_in_raw_obj_210518
length(mito_in_raw_obj_210622 )

#create a column "percent.mito" in the metadata that contains the percentage of mitochondrial gene UMI counts in each spot
all_sections_subset_QC$percent.mito <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[mito_in_raw_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.mito" in the metadata that contains the UMI counts of mitochondrial genes in each spot
all_sections_subset_QC$counts.mito <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[mito_in_raw_obj_210622, ])

```

## Protein coding content

``` {r get bmalayi protein coding gene list}

#load in wBm genes
bmalayi_protein_coding_gene_list <- readRDS("all_protein_coding_genes/bmalayi_protein_coding_genes.rds") #refhook = NULL
length(bmalayi_protein_coding_gene_list)

```

``` {r add protein coding info to metadata}

#get the list of genes in the object that intersect/match the genes in the bmalayi_protein_coding_gene_list  list
bmalayi_protein_coding_in_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_protein_coding_gene_list))
#wBm_in_obj_210518
length(bmalayi_protein_coding_in_obj_210622)

#create a column "percent.wBm" in the metadata that contains the percentage of bmalayi_protein_coding gene UMI counts in each spot
all_sections_subset_QC$percent.protein_coding <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_in_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.wBm" in the metadata that contains the UMI counts of bmalayi_protein_coding genes in each spot
all_sections_subset_QC$counts.protein_coding <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_in_obj_210622, ])

all_sections_subset_QC@meta.data
```

Protein coding - no ribo, no mito, no ribo+no mito
``` {r add protein coding info to metadata 2}

#load protein coding no ribo
bmalayi_protein_coding_no_ribo_gene_list <- readRDS("bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.rds")
length(bmalayi_protein_coding_no_ribo_gene_list)

#load protein coding no mito
bmalayi_protein_coding_no_mito_gene_list <- readRDS("bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.rds")
length(bmalayi_protein_coding_no_mito_gene_list)

#load protein coding no ribo+no mito
bmalayi_protein_coding_no_ribo_no_mito_gene_list <- readRDS("bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")
length(bmalayi_protein_coding_no_ribo_no_mito_gene_list)


```

``` {r add protein coding no ribo info to metadata 3}

#get the list of genes in the object that intersect/match the genes in the protein coding no ribo list
bmalayi_protein_coding_no_ribo_in_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_protein_coding_no_ribo_gene_list))
#wBm_in_obj_210518
length(bmalayi_protein_coding_no_ribo_in_obj_210622)

#create a column "percent.wBm" in the metadata that contains the percentage of protein coding no ribo gene UMI counts in each spot
all_sections_subset_QC$percent.protein_coding_no_ribo <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_ribo_in_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.wBm" in the metadata that contains the UMI counts of protein coding no ribo genes in each spot
all_sections_subset_QC$counts.protein_coding_no_ribo <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_ribo_in_obj_210622, ])

all_sections_subset_QC@meta.data
```

``` {r add protein coding no mtio info to metadata 4}

#get the list of genes in the object that intersect/match the genes in the protein coding no mtio list
bmalayi_protein_coding_no_mito_in_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_protein_coding_no_mito_gene_list))
#wBm_in_obj_210518
length(bmalayi_protein_coding_no_mito_in_obj_210622)

#create a column "percent.wBm" in the metadata that contains the percentage of protein coding no mtio gene UMI counts in each spot
all_sections_subset_QC$percent.protein_coding_no_mito <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_mito_in_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.wBm" in the metadata that contains the UMI counts of protein coding no mtio genes in each spot
all_sections_subset_QC$counts.protein_coding_no_mito <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_mito_in_obj_210622, ])

all_sections_subset_QC@meta.data
```

``` {r add protein coding no ribo+no mito info to metadata 5}

#get the list of genes in the object that intersect/match the genes in the protein coding no ribo+no mitolist
bmalayi_protein_coding_no_ribo_no_mito_in_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),bmalayi_protein_coding_no_ribo_no_mito_gene_list))
#wBm_in_obj_210518
length(bmalayi_protein_coding_no_ribo_no_mito_in_obj_210622)

#create a column "percent.wBm" in the metadata that contains the percentage of protein coding no ribo+no mito gene UMI counts in each spot
all_sections_subset_QC$percent.protein_coding_no_ribo_no_mito <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_ribo_no_mito_in_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.wBm" in the metadata that contains the UMI counts of protein coding no ribo+no mito genes in each spot
all_sections_subset_QC$counts.protein_coding_no_ribo_no_mito <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[bmalayi_protein_coding_no_ribo_no_mito_in_obj_210622, ])

all_sections_subset_QC@meta.data
```

## Wolbachia (wBm) content

``` {r get wBm gene list}

#load in wBm genes
wBM_all_gene_list <- readRDS("wBm_all_genes/all_wBm_genes.rds")
length(wBM_all_gene_list)

all_sections_subset_QC_wBm <- intersect(rownames(all_sections_subset_QC), wBM_all_gene_list)
all_sections_subset_QC_wBm
```

``` {r add wBm info to metadata}

#get the list of genes in the object that intersect/match the genes in the wBm list
wBm_in_obj_210622 <- Reduce(intersect, list(rownames(all_sections_subset_QC@assays$RNA),wBM_all_gene_list))
#wBm_in_obj_210518
length(wBm_in_obj_210622)

#create a column "percent.wBm" in the metadata that contains the percentage of wBm gene UMI counts in each spot
all_sections_subset_QC$percent.wBm <- (Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[wBm_in_obj_210622, ])/Matrix::colSums(all_sections_subset_QC@assays$RNA@counts))*100

#create a column "count.wBm" in the metadata that contains the UMI counts of wBm genes in each spot
all_sections_subset_QC$counts.wBm <- Matrix::colSums(all_sections_subset_QC@assays$RNA@counts[wBm_in_obj_210622, ])

all_sections_subset_QC@meta.data
```

## Save pre-filtered object

``` {r save labeled+metadata object}

saveRDS(all_sections_subset_QC, file = "bmalayi.pre.filtered.rds")

```

## Save metadata

``` {r save meta.data}

write.csv(all_sections_subset_QC@meta.data[,], file = "V10F24_041_extended_metadata.csv")

write.xlsx(all_sections_subset_QC@meta.data[,], file = "V10F24_041_extended_metadata.xlsx", col.names=TRUE, row.names=TRUE)

```

Output is the metadata, contained in the rds object, used in the remaining analysis script and the brugia count matrix and metadata rds object used in the next script, 4_quality_check_filtering.Rmd.