---
title: "QC and filtering"
author: "Hailey Sounart"
date: "8/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
   library(STutility)
  library(hdf5r)
  library(imager)
  library(akima)
  library(plotly)
  library(raster)
  library(dplyr)
  library(magrittr)
  library(ggplot2)
  library(imager) 
  library(magrittr)
  library(RColorBrewer)
  library(stringr)
  library(openxlsx)
  library(hash)
  library(readxl)
  library(purrr)
  library(harmony)
  library(DESeq2)
  library(performance)
  library(Hmisc)
  library(corrplot)
  library(pheatmap)
  library(performance)
  library(cowplot)
  library(ggpubr)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")
``` 

## Load pre-filtered object, contains metadata

## Read in Object

``` {r read in labeled object}

all_section_QC <- readRDS("bmalayi.pre.filtered.rds")
all_section_QC
```

## Remove Wolbachia (wBm) from count matrix

``` {r get wBm gene list}

#load in wBm genes
wBm_all_gene_list <- readRDS("wBm_all_genes/all_wBm_genes.rds") #refhook = NULL
length(wBm_all_gene_list)

all_section_QC_wBm <- intersect(rownames(all_section_QC), wBm_all_gene_list)
all_section_QC_wBm

all_section_QC_wBm_obj <- SubsetSTData(all_section_QC, features = all_section_QC_wBm)
all_section_QC_wBm_obj

```

``` {r get bmalayi genes}

#load in wBm genes
bmalayi_all_gene_list <- readRDS("bmalayi_all_genes/bmalayi_genes.rds")
length(bmalayi_all_gene_list)

all_section_QC_bmalayi <- intersect(rownames(all_section_QC), bmalayi_all_gene_list)
length(all_section_QC_bmalayi)

all_section_QC_bmalayi_obj <- SubsetSTData(all_section_QC, features = all_section_QC_bmalayi)
all_section_QC_bmalayi_obj


```


## Remove genes with 0 counts

``` {r remove genes}

#min.gene.count
nUMI_filt = Matrix::rowSums(all_section_QC_bmalayi_obj@assays$RNA@counts)
gene_umi_count <- rownames(all_section_QC_bmalayi_obj@assays$RNA[(nUMI_filt >= 1),])

all_section_QC_bmalayi_obj.gene.filt <- SubsetSTData(all_section_QC_bmalayi_obj, features = gene_umi_count)

cat("Genes removed: ", nrow(all_section_QC_bmalayi_obj) - nrow(all_section_QC_bmalayi_obj.gene.filt), "\n")
cat("Genes remaining: ", nrow(all_section_QC_bmalayi_obj.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(all_section_QC_bmalayi_obj.gene.filt)/nrow(all_section_QC_bmalayi_obj))*100), "\n")

all_section_QC_bmalayi_obj.gene.filt

```

Check metadata
``` {r check metadata}

all_section_QC_bmalayi_obj.gene.filt@meta.data

```

## Violin plots - per sample

```{r violin plots sample}

sample_levels <- c("BM1","BM2","BM3")
all_section_QC_bmalayi_obj.gene.filt@meta.data$worm_sample <- factor(x = all_section_QC_bmalayi_obj.gene.filt@meta.data$worm_sample, levels = sample_levels)

``` 


```{r violin plots sample 2}

fill <- c("#AEC7E8","#009E73","#0072B2")

bmalayi_vlnplot_unique_gene_sample_no_pt <- VlnPlot(all_section_QC_bmalayi_obj.gene.filt, features = c("nFeature_RNA"), split.by = "worm_sample", group.by = "worm_sample", cols = fill, pt.size = 0)
bmalayi_vlnplot_umi_sample_no_pt  <- VlnPlot(all_section_QC_bmalayi_obj.gene.filt, features = c("nCount_RNA"), split.by = "worm_sample", group.by = "worm_sample", cols = fill, pt.size = 0)

``` 

```{r violin plots sample 3}

dir.create(paste0("qc_plots"))

ggsave(file = "qc_plots/bmalayi_violin_plot_umis_raw_ctrl_no_pt.pdf", plot = bmalayi_vlnplot_umi_sample_no_pt, dpi = 300)
ggsave(file = "qc_plots/bmalayi_violin_plot_genes_raw_ctrl_no_pt.pdf", plot = bmalayi_vlnplot_unique_gene_sample_no_pt, dpi = 300)

``` 

## Violin plots - per section

```{r violin plots section}
unique(all_section_QC_bmalayi_obj.gene.filt@meta.data$sample_section)
length(unique(all_section_QC_bmalayi_obj.gene.filt@meta.data$sample_section))
section_levels <- c("BM1_1","BM1_2","BM1_3","BM1_4","BM1_5","BM1_6","BM1_7","BM1_8","BM1_9","BM2_2","BM2_3","BM2_4","BM2_5","BM2_6","BM2_7","BM2_8","BM2_9","BM2_10","BM2_11","BM2_12","BM3_1","BM3_2","BM3_3","BM3_4","BM3_7","BM3_8","BM3_9","BM3_10","BM3_11","BM3_12")
all_section_QC_bmalayi_obj.gene.filt@meta.data$sample_section <- factor(x = all_section_QC_bmalayi_obj.gene.filt@meta.data$sample_section, levels = section_levels)

``` 

```{r violin plots section 2}

fill_3 <- c("#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#AEC7E8","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#009E73","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2","#0072B2")


bmalayi_vlnplot_unique_gene_section_no_pt <- VlnPlot(all_section_QC_bmalayi_obj.gene.filt, features = c("nFeature_RNA"), split.by = "sample_section", group.by = "sample_section", cols = fill_3, pt.size = 0)
bmalayi_vlnplot_umi_section_no_pt  <- VlnPlot(all_section_QC_bmalayi_obj.gene.filt, features = c("nCount_RNA"), split.by = "sample_section", group.by = "sample_section", cols = fill_3, pt.size = 0)

bmalayi_vlnplot_unique_gene_section_no_pt
bmalayi_vlnplot_umi_section_no_pt

``` 

```{r violin plots section 3}

ggsave(file = "qc_plots/bmalayi_violin_plot_umis_raw_ctr_sectionl_no_pt.pdf", plot = bmalayi_vlnplot_umi_section_no_pt, dpi = 300, height = 400, width = 600, units = "mm")
ggsave(file = "qc_plots/bmalayi_violin_plot_genes_raw_ctrl_section_no_pt.pdf", plot = bmalayi_vlnplot_unique_gene_section_no_pt, dpi = 300, height = 400, width = 600, units = "mm")

``` 

## Correlation plots - per sample

```{r corr plots sample}


umis.all.avg <- all_section_QC_bmalayi_obj.gene.filt@assays$RNA@counts
sum.expr.all.avg <- lapply(unique(all_section_QC_bmalayi_obj.gene.filt$worm_sample), function(treat) {
  umis_subset.all.avg <- umis.all.avg[, all_section_QC_bmalayi_obj.gene.filt$worm_sample %in% treat]  #subset counts dataframe
  print(class(umis_subset.all.avg))
  log1p(Matrix::rowMeans(umis_subset.all.avg)) #log1p(Matrix::rowMeans(umis_subset2))
})

sum.expr.all.avg <- as.data.frame(sum.expr.all.avg)
colnames(sum.expr.all.avg) <- unique(all_section_QC_bmalayi_obj.gene.filt$worm_sample)
worm_sample_raw_unfil_df <- sum.expr.all.avg
worm_sample_raw_unfil_df
```

Control worm sample comparisons
```{r corr plots sample 2}


colnames(worm_sample_raw_unfil_df) <- c("BM1", "BM2","BM3")
worm_sample_raw_unfil_df


p1 <- ggplot(worm_sample_raw_unfil_df, aes(BM1, BM2)) + geom_point(alpha = 1,colour = "#009E73") + stat_cor(method = "pearson", label.x = 0, label.y = 3.7) + geom_smooth(method = "lm", fullrange=TRUE, color="#009E73", se = TRUE, level = 0.95)  + theme_minimal() + labs(title="Average gene expression between samples",
       x="BM1 (log (1 + avg raw UMI counts per spot))", y = "BM2 (log (1 + avg raw UMI counts per spot))")#+ ggtitle("Blah") 

p2 <- ggplot(worm_sample_raw_unfil_df, aes(BM1, BM3)) + geom_point(colour = "#009E73",alpha = 1) + stat_cor(method = "pearson", label.x = 0, label.y = 3.7) + geom_smooth(method = "lm", fullrange=TRUE, color="#009E73", se = TRUE, level = 0.95) + theme_minimal() + labs(title="Average gene expression between samples",
       x="BM1 (log (1 + avg raw UMI counts per spot))", y = "BM3 (log (1 + avg raw UMI counts per spot))")

p3 <- ggplot(worm_sample_raw_unfil_df, aes(BM2, BM3)) + geom_point(colour = "#009E73",alpha = 1) + stat_cor(method = "pearson", label.x = 0, label.y = 3.7) + geom_smooth(method = "lm", fullrange=TRUE, color="#009E73", se = TRUE, level = 0.95) + theme_minimal() + labs(title="Average gene expression between samples",
       x="BM2 (log (1 + avg raw UMI counts per spot))", y = "BM3 (log (1 + avg raw UMI counts per spot))")


ggsave(file = "qc_plots/bmalayi_corr_worm_raw_unfilt_1_2.pdf", plot = p1, dpi = 300)
ggsave(file = "qc_plots/bmalayi_corr_worm_raw_unfilt_1_3.pdf", plot = p2, dpi = 300)
ggsave(file = "qc_plots/bmalayi_corr_worm_raw_unfilt_2_3.pdf", plot = p3, dpi = 300)


``` 

## Correlation plots - per section

```{r corr plots section}


umis.all.avg <- all_section_QC_bmalayi_obj.gene.filt@assays$RNA@counts
sum.expr.all.avg <- lapply(unique(all_section_QC_bmalayi_obj.gene.filt$sample_section), function(treat) {
  umis_subset.all.avg <- umis.all.avg[, all_section_QC_bmalayi_obj.gene.filt$sample_section %in% treat]  #subset counts dataframe
  print(class(umis_subset.all.avg))
  log1p(Matrix::rowMeans(umis_subset.all.avg))
})

sum.expr.all.avg <- as.data.frame(sum.expr.all.avg)
colnames(sum.expr.all.avg) <- unique(all_section_QC_bmalayi_obj.gene.filt$sample_section)
worm_sample_raw_unfil_df <- sum.expr.all.avg
worm_sample_raw_unfil_df
```

Control worm section comparisons - all worms
```{r corr plots section 2}

dir.create(paste0("qc_plots/sections"))

colnames(worm_sample_raw_unfil_df) <- c("BM1_1","BM1_2","BM1_3","BM1_4","BM1_5","BM1_6","BM1_7","BM1_8","BM1_9","BM2_2","BM2_3","BM2_4","BM2_5","BM2_6","BM2_7","BM2_8","BM2_9","BM2_10","BM2_11","BM2_12","BM3_1","BM3_2","BM3_3","BM3_4","BM3_7","BM3_8","BM3_9","BM3_10","BM3_11","BM3_12")
worm_sample_raw_unfil_df

for (i in colnames(worm_sample_raw_unfil_df)) {
  first_section <- i
  first_section <- noquote(first_section)
  print(first_section)

  for (j in colnames(worm_sample_raw_unfil_df)){ 
    
    second_section <- noquote(j)
    print(second_section)
    
    p <- aes_(as.name(first_section), as.name(second_section))
    print(p)
    
    p <- ggplot(worm_sample_raw_unfil_df, aes_(as.name(first_section),as.name(second_section))) + geom_point(alpha = 1,colour = "#009E73") + stat_cor(method = "pearson", label.x = 0, label.y = 3.7) + geom_smooth(method = "lm", fullrange=TRUE, color="#009E73", se = TRUE, level = 0.95)  + theme_minimal() + labs(title="Average gene expression between sections",           x=paste0(first_section, " (log (1 + avg raw UMI counts per spot))"), y = paste0(second_section, " (log (1 + avg raw UMI counts per spot))"))

    ggsave(file = paste0("qc_plots/sections/bmalayi_corr_worm_raw_unfilt_", first_section, "_", second_section, ".pdf"), plot = p, dpi = 300)
  }
}

``` 

## Load protein coding genes, no ribo, no mito
```{r load protein genes}

bmalayi_protein_coding.no.ribo.no.mito <- readRDS("bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")
length(bmalayi_protein_coding.no.ribo.no.mito)

``` 

## Keep only protein coding genes (no mito and no ribo)

```{r keep only protein coding genes}

all_section_QC_bmalayi_obj.gene.filt

all_section_QC_bmalayi_obj.protein.coding.no.ribo.no.mito <- SubsetSTData(all_section_QC_bmalayi_obj.gene.filt, features = rownames(all_section_QC_bmalayi_obj.gene.filt)[(rownames(all_section_QC_bmalayi_obj.gene.filt) %in% bmalayi_protein_coding.no.ribo.no.mito)])

all_section_QC_bmalayi_obj.protein.coding.no.ribo.no.mito

``` 

## Filter spots 

Filter low quality spots
```{r filter spots}

bmalayi_obj.spot.filt <- SubsetSTData(all_section_QC_bmalayi_obj.protein.coding.no.ribo.no.mito, expression = nFeature_RNA >= 30 & nCount_RNA >= 50 & percent.mito < 3 & percent.ribo < 30 & nCount_RNA < 10000) #& percent.mito < 10 & percent.ribo < 40
cat("Spots removed: ", ncol(all_section_QC_bmalayi_obj.protein.coding.no.ribo.no.mito) - ncol(bmalayi_obj.spot.filt), "\n")
cat("Spots remaining: ", ncol(bmalayi_obj.spot.filt), "\n")
cat("Percent Spots remaining: ", (ncol(bmalayi_obj.spot.filt)/ncol(all_section_QC_bmalayi_obj.protein.coding.no.ribo.no.mito)*100), "\n")

bmalayi_obj.spot.filt

```

## Filter genes

Filter based on number of spots per gene & total UMI counts per gene.
```{r filter genes}

#min.gene.spots
nSpots_filt = Matrix::rowSums(bmalayi_obj.spot.filt@assays$RNA@counts > 0)
gene_spot_count <- rownames(bmalayi_obj.spot.filt@assays$RNA[(nSpots_filt >= 1),])
  
#min.gene.count
nUMI_filt = Matrix::rowSums(bmalayi_obj.spot.filt@assays$RNA@counts)
gene_umi_count <- rownames(bmalayi_obj.spot.filt@assays$RNA[(nUMI_filt >= 1),])

bmalayi_obj.spot.gene.filt <- SubsetSTData(bmalayi_obj.spot.filt, features = gene_umi_count)
bmalayi_obj.spot.gene.filt <-SubsetSTData(bmalayi_obj.spot.gene.filt, features = gene_spot_count)

cat("Genes removed: ", nrow(bmalayi_obj.spot.filt) - nrow(bmalayi_obj.spot.gene.filt), "\n")
cat("Genes remaining: ", nrow(bmalayi_obj.spot.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(bmalayi_obj.spot.gene.filt)/nrow(bmalayi_obj.spot.filt))*100), "\n")


bmalayi_obj.spot.gene.filt

```

## Re-calc metadata

```{r load gene lists, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

#load in mitochondrial gene list
bmalayi_mito_gene_list <- readRDS("mitochondrial_genes/bmalayi_mito_genes.rds")
bmalayi_mito_gene_list

#load in rRNA genes
bmalayi_rRNA_gene_list <- readRDS("rRNA_genes/bmalayi_rRNA_genes.rds")

#load in ribosomal protein coding genes
bmalayi_ribo_gene_list <- readRDS("ribosomal_genes/bmalayi_ribo_genes.rds")
#bmalayi_ribo_gene_list

#combine rRNA and ribosomal protein coding genes into a single list
bmalayi_all_ribo_list <- c(bmalayi_rRNA_gene_list, bmalayi_ribo_gene_list)

#load in bmalayi protein coding 
bmalayi_coding_list <- readRDS("all_protein_coding_genes/bmalayi_protein_coding_genes.rds")

#load in bmalayi protein coding, no ribo
bmalayi_coding_no_ribo_list <- readRDS("bmalayi_protein_coding_no_ribo/bmalayi_protein_coding.no.ribo.rds")

#load in bmalayi protein coding, no mito
bmalayi_coding_no_mito_list <- readRDS("bmalayi_protein_coding_no_mito/bmalayi_protein_coding.no.mito.rds")

#load in bmalayi protein coding,no ribo and no mito
bmalayi_coding_no_ribo_no_mito_list <- readRDS("bmalayi_protein_coding_no_ribo_no_mito/bmalayi_protein_coding.no.ribo.no.mito.rds")

```

```{r recalc metadata, all spots 1, fig.width = 10, fig.height = 5, out.height = "100%", out.width = "100%"}

print(sum(bmalayi_obj.spot.gene.filt$percent.mito))
print(sum(bmalayi_obj.spot.gene.filt$counts.mito))

bmalayi_obj.spot.gene.filt.new <- bmalayi_obj.spot.gene.filt

mito_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_mito_gene_list))
bmalayi_obj.spot.gene.filt.new$percent.mito <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[mito_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.mito <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[mito_in_obj, ])
rRNA_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_rRNA_gene_list))
bmalayi_obj.spot.gene.filt.new$percent.rRNA <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[rRNA_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.rRNA <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[rRNA_in_obj, ])
ribo_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_all_ribo_list))
bmalayi_obj.spot.gene.filt.new$percent.ribo <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[ribo_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.ribo <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[ribo_in_obj, ])
coding_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_coding_list))
bmalayi_obj.spot.gene.filt.new$percent.protein_coding <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[coding_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.protein_coding <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[coding_in_obj, ])
coding_no_ribo_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_coding_no_ribo_list))
bmalayi_obj.spot.gene.filt.new$percent.protein_coding_no_ribo <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[ coding_no_ribo_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.protein_coding_no_ribo <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[ coding_no_ribo_in_obj, ])
coding_no_mito_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_coding_no_mito_list))
bmalayi_obj.spot.gene.filt.new$percent.protein_coding_no_mito <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[coding_no_mito_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.protein_coding_no_mito <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[coding_no_mito_in_obj, ])
coding_no_ribo_no_mito_in_obj <- Reduce(intersect, list(rownames(bmalayi_obj.spot.gene.filt@assays$RNA),bmalayi_coding_no_ribo_no_mito_list))
bmalayi_obj.spot.gene.filt.new$percent.protein_coding_no_ribo_no_mito <- (Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts[coding_no_ribo_no_mito_in_obj, ])/Matrix::colSums(bmalayi_obj.spot.gene.filt@assays$RNA@counts))*100
bmalayi_obj.spot.gene.filt.new$counts.protein_coding_no_ribo_no_mito <- Matrix::colSums(bmalayi_obj.spot.gene.filt.new@assays$RNA@counts[coding_no_ribo_no_mito_in_obj, ])


print(sum(bmalayi_obj.spot.gene.filt.new$percent.mito))
print(sum(bmalayi_obj.spot.gene.filt.new$counts.mito))
bmalayi_obj.spot.gene.filt.new


```

## Save spot, gene filt objects

``` {r save filtered bmalayi object}

saveRDS(bmalayi_obj.spot.gene.filt.new, file = "bmalay_filt_obj.rds")

```

## Wolbachia processing

Subset Wolbachia object to only include those spots in the brugia object

```{r filter wBm obj}

unique(all_section_QC_wBm_obj$worm_sample)

bmalayi_obj.spot.gene.filt.spots <- colnames(bmalayi_obj.spot.gene.filt.new@assays$RNA[,])
length(bmalayi_obj.spot.gene.filt.spots)
#head(bmalayi_obj.spot.gene.filt.spots)

all_section_QC_wBm_obj.brugia.filt.spots <- SubsetSTData(all_section_QC_wBm_obj, spots = bmalayi_obj.spot.gene.filt.spots)

all_section_QC_wBm_obj.brugia.filt.spots

```

Remove genes with 0 summed UMIs - from R objects
```{r remove 0 count genes}

#all_section_QC_wBm_obj), rownames(wBm_ctrl_only_no_C10

#min.gene.spots
nSpots_filt = Matrix::rowSums(all_section_QC_wBm_obj.brugia.filt.spots@assays$RNA@counts > 0)
gene_spot_count <- rownames(all_section_QC_wBm_obj.brugia.filt.spots@assays$RNA[(nSpots_filt >= 1),])
  
#min.gene.count
nUMI_filt = Matrix::rowSums(all_section_QC_wBm_obj.brugia.filt.spots@assays$RNA@counts)
gene_umi_count <- rownames(all_section_QC_wBm_obj.brugia.filt.spots@assays$RNA[(nUMI_filt >= 1),])

all_section_QC_wBm_obj.brugia.filt.spots.gene.filt <- SubsetSTData(all_section_QC_wBm_obj.brugia.filt.spots, features = gene_umi_count)
all_section_QC_wBm_obj.brugia.filt.spots.gene.filt <-SubsetSTData(all_section_QC_wBm_obj.brugia.filt.spots.gene.filt, features = gene_spot_count)

cat("Genes removed: ", nrow(all_section_QC_wBm_obj.brugia.filt.spots) - nrow(all_section_QC_wBm_obj.brugia.filt.spots.gene.filt), "\n")
cat("Genes remaining: ", nrow(all_section_QC_wBm_obj.brugia.filt.spots.gene.filt), "\n")
cat("Percent genes remaining: ", ((nrow(all_section_QC_wBm_obj.brugia.filt.spots.gene.filt)/nrow(all_section_QC_wBm_obj.brugia.filt.spots))*100), "\n")


all_section_QC_wBm_obj.brugia.filt.spots.gene.filt


``` 

## Save Wolbachia objects

```{r save wBm obj}

saveRDS(all_section_QC_wBm_obj.brugia.filt.spots.gene.filt, file = "wBm.filt.rds")

```


Outputs are the violin plots and correlation plots shown in Figure 1 of the main manuscript, the filtered brugia count matrix and metadata rds object used in the next script, 5_normalization.Rmd, and the Wolbachia (wBm) filtered count matrix and metadata rds object.