---
title: "Clustering and Differential Expression analysis"
author: "Hailey Sounart"
date: "8/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(hdf5r)
  library(imager)
  library(akima)
  library(plotly)
  library(raster)
  library(dplyr)
  library(magrittr)
  library(ggplot2)
  library(imager) 
  library(magrittr)
  library(RColorBrewer)
  library(stringr)
  library(openxlsx)
  library(hash)
  library(readxl)
  library(purrr)
  library(harmony)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")
``` 

## Read in bmalayi normalized data

``` {r read normalized object}

bmalayi_sctransform_per_sample_norm_obj <- readRDS("bmalay_filt_obj.list.nfeat.merged.rds")
bmalayi_sctransform_per_sample_norm_obj

```

## Load integration features
``` {r load integ feats}

bmalayi_sctransform_per_sample_norm_obj_int_feats <- readRDS("bmalay_filt_obj.list.nfeat.integration.features.rds")
length(bmalayi_sctransform_per_sample_norm_obj_int_feats)

```

## Dimensionality Reduction

PCA
``` {r dim reduction}


bmalayi_sctransform_per_sample_norm_obj.PCA <- RunPCA(object = bmalayi_sctransform_per_sample_norm_obj, assay = "SCT", features = bmalayi_sctransform_per_sample_norm_obj_int_feats)

```

## Harmony (data integration and batch effect removal) if needed

```{r run harmony PCA}

set.seed(23)
bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- RunHarmony(object = bmalayi_sctransform_per_sample_norm_obj.PCA, vars_use = c("worm_sample"), group.by.vars = c("worm_sample"), reduction = "pca", assay.use = "SCT", plot_convergence = TRUE, dims.use = 1:10, theta = 0, max.iter.harmony = 80)

``` 

## Clustering 

``` {r clustering, fig.width = 4, fig.height = 4, out.height = "100%", out.width = "100%"}

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony  <- FindNeighbors(object = bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, verbose = FALSE, reduction = "harmony", dims = 1:10, k.param = 10) 

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- FindClusters(object = bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, verbose = FALSE, resolution = 0.3) 

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- RunUMAP(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, reduction = "harmony", dims = 1:10)

```


## View Clusters

```{r view clusters}

fill=c("#41ab5d","#fc9272","#4292c6","#df65b0")

DimPlot(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, label = FALSE, cols = fill, pt.size = 3)

```
## Save PCA, harmony object

``` {r save clustered object}

saveRDS(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, file = "bmalayi_clusters.rds")

```

## DE analysis - marker genes per cluster

``` {r find cluster markers}

markers_per_clu <- FindAllMarkers(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, assay = "SCT", logfc.threshold = 0.1)
markers_per_clu

```

## p val filt

``` {r filt cluster markers}

markers_per_clu_filt <- markers_per_clu[(markers_per_clu$p_val_adj < 0.05),]
markers_per_clu_filt

```

## combined into DE genes into 1 table

Separate each cluster markers by cluster
``` {r split cluster markers}

markers_per_clu_filt_list <- split(markers_per_clu_filt, markers_per_clu_filt$cluster)
markers_per_clu_filt_list

```

Rename genes
``` {r rename genes}
markers_per_clu_filt_list_rename <- lapply(markers_per_clu_filt_list, function(i) {
  rownames(i) <- i$gene
  i
})
markers_per_clu_filt_list_rename

```

```{r merge dataframes 1}

markers_per_clu_filt_list_1_2_rename_merged <- transform(merge(markers_per_clu_filt_list_rename[[1]], markers_per_clu_filt_list_rename[[2]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
markers_per_clu_filt_list_1_2_rename_merged

``` 

```{r merge dataframes 2}

markers_per_clu_filt_list_1_2_3_rename_merged <- transform(merge(markers_per_clu_filt_list_1_2_rename_merged, markers_per_clu_filt_list_rename[[3]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
markers_per_clu_filt_list_1_2_3_rename_merged

``` 

```{r merge dataframes 3}

markers_per_clu_filt_list_rename_merged <- transform(merge(markers_per_clu_filt_list_1_2_3_rename_merged, markers_per_clu_filt_list_rename[[4]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
markers_per_clu_filt_list_rename_merged

``` 

## Save cluster marker genes

``` {r save cluster de marker genes}

write.xlsx(markers_per_clu_filt_list_rename_merged , file = "clus_de_genes_p0.05.xlsx", col.names=TRUE, row.names=TRUE)
```
