---
title: "Normalization"
author: "Hailey Sounart"
date: "8/16/2022"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
```

```{r set working directory}

#setwd("<path to working directory where count matrices for all samples reside>")
setwd("/Users/hailey.sounart/Documents/brugia_malayi/230911_final_code/")
``` 

## Read in Filtered Seurat Objects

Read in filtered object
``` {r read filt object}

bmalay_filt_obj <- readRDS("bmalay_filt_obj.rds")
bmalay_filt_obj

```

# Normalize samples separately

First split each worm_sample to be is own object and create a list of worm sample objects

```{r separate worm samples}

bmalay_filt_obj.unique.names <- unique(bmalay_filt_obj$worm_sample)
bmalay_filt_obj.unique.names

bmalay_filt_obj.list.nfeat <- lapply(bmalay_filt_obj.unique.names, function(i){SubsetSTData(bmalay_filt_obj, expression = (worm_sample == i))})

bmalay_filt_obj.list.nfeat
```

```{r normalize each sample object separately}

for (i in 1:length(bmalay_filt_obj.list.nfeat)) {
  bmalay_filt_obj.list.nfeat[[i]] <- SCTransform(bmalay_filt_obj.list.nfeat[[i]], vars.to.regress = "nFeature_RNA", return.only.var.genes = FALSE)
}

```

## Integration Features

```{r integrate}

bmalay_filt_obj.list.nfeat.integration.features <- SelectIntegrationFeatures(object.list = bmalay_filt_obj.list.nfeat, nfeatures = 7000)
saveRDS(bmalay_filt_obj.list.nfeat.integration.features, file = "bmalay_filt_obj.list.nfeat.integration.features.rds")
length(bmalay_filt_obj.list.nfeat.integration.features)

bmalay_filt_obj.list.nfeat.prepsctint <- PrepSCTIntegration(object.list = bmalay_filt_obj.list.nfeat, anchor.features = bmalay_filt_obj.list.nfeat.integration.features, verbose = FALSE)

saveRDS(bmalay_filt_obj.list.nfeat.prepsctint, file = "bmalay_filt_obj.list.nfeat.prepsctint.rds")

```

## Merge samples
``` {r merge normalized objects}

bmalay_filt_obj.list.nfeat.merged <- MergeSTData(x = bmalay_filt_obj.list.nfeat.prepsctint[[1]], y = bmalay_filt_obj.list.nfeat.prepsctint[2:length(bmalay_filt_obj.list.nfeat.prepsctint)], merge.data = TRUE)

bmalay_filt_obj.list.nfeat.merged@meta.data
bmalay_filt_obj.list.nfeat.merged

```

## Save Normalized object
```{r save normalized object}

#save control normalized data, merged
bmalay_filt_obj.list.nfeat.merged

saveRDS(bmalay_filt_obj.list.nfeat.merged, file = "bmalay_filt_obj.list.nfeat.merged.rds")

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

The outputs are the integration features from SelectIntegrationFeatures, output of PrepSCTIntegration, and the normalized rds object of brugia count matrix and metadata. The integration features and normalized rds object of brugia count matrix and metadata are used in the next scripts, 6_clustering_analysis.Rmd, 7_co_localization_analysis.Rmd, and in the 3d_model_project.zip for making the 3D model plots show in Figure 2A, Figure 3B-F, Figure 4B-D, and Figure 5C-D.
