---
title: "Clustering and Differential Expression analysis"
author: "Hailey Sounart"
date: "8/16/2022"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(dplyr)
  library(ggplot2)
  library(openxlsx)
  library(harmony)
```

```{r set working directory}

#setwd("<path to working directory where count matrices for all samples reside>")
setwd("/Users/hailey.sounart/Documents/brugia_malayi/230911_final_code/")
``` 

## Read in bmalayi normalized data

``` {r read normalized object}

bmalayi_sctransform_per_sample_norm_obj <- readRDS("bmalay_filt_obj.list.nfeat.merged.rds")
bmalayi_sctransform_per_sample_norm_obj

```

## Load integration features
``` {r load integ feats}

bmalayi_sctransform_per_sample_norm_obj_int_feats <- readRDS("bmalay_filt_obj.list.nfeat.integration.features.rds")
length(bmalayi_sctransform_per_sample_norm_obj_int_feats)

```

## Dimensionality Reduction

PCA
``` {r dim reduction}

bmalayi_sctransform_per_sample_norm_obj.PCA <- RunPCA(object = bmalayi_sctransform_per_sample_norm_obj, assay = "SCT", features = bmalayi_sctransform_per_sample_norm_obj_int_feats)

```

## Harmony (data integration and batch effect removal) if needed

```{r run harmony PCA}

set.seed(23)
bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- RunHarmony(object = bmalayi_sctransform_per_sample_norm_obj.PCA, vars_use = c("worm_sample"), group.by.vars = c("worm_sample"), reduction = "pca", assay.use = "SCT", plot_convergence = TRUE, dims.use = 1:10, theta = 0, max.iter.harmony = 80)

``` 

## Clustering 

``` {r clustering, fig.width = 4, fig.height = 4, out.height = "100%", out.width = "100%"}

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony  <- FindNeighbors(object = bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, verbose = FALSE, reduction = "harmony", dims = 1:10, k.param = 10) 

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- FindClusters(object = bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, verbose = FALSE, resolution = 0.3) 

bmalayi_sctransform_per_sample_norm_obj.PCA.harmony <- RunUMAP(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, reduction = "harmony", dims = 1:10)

```

## View Clusters

```{r view clusters}

# Figure 2B

fill=c("#41ab5d","#fc9272","#4292c6","#dd648f")

DimPlot(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, label = FALSE, cols = fill, pt.size = 3)

```

## Save PCA, harmony object

``` {r save clustered object}

saveRDS(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, file = "bmalayi_clusters.rds")

```

## DE analysis - marker genes per cluster

``` {r find cluster markers}

markers_per_clu <- FindAllMarkers(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony, assay = "SCT", logfc.threshold = 0.1)
markers_per_clu

```

## p val filt

``` {r filt cluster markers}

#filter for less than p-value adjusted < 0.05
markers_per_clu_filt <- markers_per_clu[(markers_per_clu$p_val_adj < 0.05),]
markers_per_clu_filt

#split for each cluster
markers_per_clu_filt_list <- split(markers_per_clu_filt, markers_per_clu_filt$cluster)
markers_per_clu_filt_list

#rename rownames
markers_per_clu_filt_list_rename <- lapply(markers_per_clu_filt_list, function(i) {
  rownames(i) <- i$gene
  i
})
markers_per_clu_filt_list_rename

```
```{r merge all filt clu dataframes}

markers_per_clu_filt_list_1_2_rename_merged <- transform(merge(markers_per_clu_filt_list_rename[[1]], markers_per_clu_filt_list_rename[[2]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL)
markers_per_clu_filt_list_1_2_rename_merged

markers_per_clu_filt_list_1_2_3_rename_merged <- transform(merge(markers_per_clu_filt_list_1_2_rename_merged, markers_per_clu_filt_list_rename[[3]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
markers_per_clu_filt_list_1_2_3_rename_merged

markers_per_clu_filt_list_rename_merged <- transform(merge(markers_per_clu_filt_list_1_2_3_rename_merged, markers_per_clu_filt_list_rename[[4]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL)
markers_per_clu_filt_list_rename_merged

``` 

```{r rename filt clu genes}

#rename genes

filt_clu_genes <- rownames(markers_per_clu_filt_list_rename_merged)

filt_clu_genes_renamed <- gsub("bmalayi-", "", filt_clu_genes)
#filt_clu_genes_renamed
```


## combined DE genes into 1 table

Rename genes
``` {r rename clu all genes}

#Rename genes

markers_per_clu
markers_per_clu_renamed <- markers_per_clu
markers_per_clu_renamed_genes <- gsub("bmalayi-", "", rownames(markers_per_clu_renamed))
rownames(markers_per_clu_renamed) <- markers_per_clu_renamed_genes
markers_per_clu_renamed

markers_per_clu_renamed$gene <- gsub("bmalayi-", "", (markers_per_clu_renamed$gene))
markers_per_clu_renamed

```


``` {r split cluster markers}

#Separate each cluster markers by cluster

markers_per_clu_list <- split(markers_per_clu_renamed, markers_per_clu_renamed$cluster)
markers_per_clu_list

```


``` {r fix clu gene names}

#Rename genes

markers_per_clu_list_rename <- lapply(markers_per_clu_list, function(i) {
  rownames(i) <- i$gene
  i
})
markers_per_clu_list_rename

```


```{r merge dataframes 1}

markers_per_clu_list_1_2_rename_merged <- transform(merge(markers_per_clu_list_rename[[1]], markers_per_clu_list_rename[[2]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL)
markers_per_clu_list_1_2_rename_merged

``` 

```{r merge dataframes 2}

markers_per_clu_list_1_2_3_rename_merged <- transform(merge(markers_per_clu_list_1_2_rename_merged, markers_per_clu_list_rename[[3]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
markers_per_clu_list_1_2_3_rename_merged

``` 

```{r merge dataframes 3}

markers_per_clu_list_rename_merged <- transform(merge(markers_per_clu_list_1_2_3_rename_merged, markers_per_clu_list_rename[[4]],by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL)
markers_per_clu_list_rename_merged

``` 

```{r edit clu markers table}

#remove columns
clu_markers_df <- markers_per_clu_list_rename_merged %>% select(-cluster.x, -gene.x, -cluster.y, -gene.y, -cluster.x.1, -gene.x.1, -cluster.y.1, -gene.y.1)

#rename columns
colnames(clu_markers_df) <- c("p_val.1","avg_logFC.1","pct.1.1","pct.2.1","p_val_adj.1","p_val.2","avg_logFC.2","pct.1.2","pct.2.2","p_val_adj.2","p_val.3","avg_logFC.3","pct.1.3","pct.2.3","p_val_adj.3","p_val.4","avg_logFC.4","pct.1.4","pct.2.4","p_val_adj.4")
clu_markers_df

```

```{r filt clu table}

#select genes with adjusted p-value < 0.05 for genes that are markers for a specific cluster
clu_markers_df_select <- clu_markers_df[(rownames(clu_markers_df) %in% filt_clu_genes_renamed),]
clu_markers_df_select
```
## Read in annotations
```{r read in bmalayi annotation}

#read in the bmalayi annotation file as a table
all_bmalayi_gene_annotations <- read.table("brugia_malayi.gene_biotype_product.map.tsv", sep="\t", header=T, stringsAsFactors = F)
all_bmalayi_gene_annotations

```

## Add anotations into table

```{r filt annot table}

filt_annot_table <- all_bmalayi_gene_annotations[(all_bmalayi_gene_annotations$gene_id %in% rownames(clu_markers_df_select)), ]
rownames(filt_annot_table) <- filt_annot_table$gene_id
filt_annot_table

```

```{r merge dataframes annot}

#Merge annotations into dataframe
clu_markers_df_select_with_annotation <- transform(merge(clu_markers_df_select, filt_annot_table,by="row.names",all.x=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
clu_markers_df_select_with_annotation
colnames(clu_markers_df_select_with_annotation)[colnames(clu_markers_df_select_with_annotation) == 'gene_id'] <- 'gene'
clu_markers_df_select_with_annotation

``` 

## Save cluster marker genes (Supplementary Data 2)

``` {r save cluster de marker genes}

#Supplementary Data 2
write.xlsx(clu_markers_df_select_with_annotation, file = "Supplementary_Data_2.xlsx", col.names=TRUE, row.names=TRUE)
```

## Plots

### Figure 2B

```{r rename clus numbers}

new_clus_names <- bmalayi_sctransform_per_sample_norm_obj.PCA.harmony
new_clus_names$seurat_clusters <- factor(bmalayi_sctransform_per_sample_norm_obj.PCA.harmony$seurat_clusters, 
         levels = c("0", "1", "2", "3"), 
         labels = c("1", "2", "3", "4"))

new_clus_names@meta.data
```

```{r save umap}

#Figure 2B

fill=c("#41ab5d","#fc9272","#4292c6","#dd648f")

clu_umap <- DimPlot(new_clus_names, group.by = "seurat_clusters", label = FALSE, cols = fill, pt.size = 3) 
clu_umap

ggsave(file = "Figure_2B.pdf", plot = clu_umap , width = 15, height =10, dpi = 300)

```

### Figure 3

``` {r text genes}

text_genes <- c("Bm7941", "Bm2751", "Bm8720", "Bm6643", "Bma-cul-2", "Bma-cyb-3","Bm13981", "Bma-gld-1", "Bma-ubq-2", "Bm13653", "Bm2195", "Bma-hsp-25", "Bma-far-1", "Bm12803", "Bma-cht-1.2", "Bm4112","Bm13693", "Bma-nmad-1", "Bm12794", "Bm5705", "Bma-hsp-17", "Bma-enpl-1", "Bm2151", "Bma-prdx-2", "Bma-rpt-4", "Bma-ned-8", "Bm11071", "Bm2854", "Bma-col-109", "Bma-pqn-35", "Bm2467", "Bma-ctg-1")
length(text_genes)

```

```{r text genes other name}

clus_text_genes <- c("WBGene00228202", "WBGene00223012", "WBGene00228981", "WBGene00226904", "WBGene00231090", "WBGene00225880", "WBGene00234242", "WBGene00226227", "WBGene00233912", "WBGene00233914", "WBGene00222456", "WBGene00222349", "WBGene00223407", "WBGene00233064", "WBGene00223435", "WBGene00224373", "WBGene00233954", "WBGene00223499", "WBGene00233055", "WBGene00225966", "WBGene00230858", "WBGene00225845", "WBGene00222412", "WBGene00223487", "WBGene00223805", "WBGene00224389", "WBGene00231332", "WBGene00223115", "WBGene00226682", "WBGene00230319", "WBGene00222728", "WBGene00224893")
length(clus_text_genes)

clus_text_genes_rename <- paste0("bmalayi-", clus_text_genes)
length(clus_text_genes_rename)

```


```{r plot dotplot clus}

#Figure 3A

text_genes_clus_dotplot <- DotPlot(new_clus_names, assay = "SCT", clus_text_genes_rename, cols = c("#feebe2", "#7a0177"), group.by = "seurat_clusters", dot.scale = 10) + RotatedAxis() + scale_x_discrete(labels = c("bmalayi-WBGene00228202" = "Bm7941", "bmalayi-WBGene00223012" = "Bm2751", "bmalayi-WBGene00228981" = "Bm8720", "bmalayi-WBGene00226904"= "Bm6643", "bmalayi-WBGene00231090" = "Bma-cul-2", "bmalayi-WBGene00225880" = "Bma-cyb-3", "bmalayi-WBGene00234242" = "Bm13981", "bmalayi-WBGene00226227" = "Bma-gld-1", "bmalayi-WBGene00233912" = "Bma-ubq-2", "bmalayi-WBGene00233914"= "Bm13653", "bmalayi-WBGene00222456" = "Bm2195", "bmalayi-WBGene00222349" = "Bma-hsp-25", "bmalayi-WBGene00223407" = "Bma-far-1", "bmalayi-WBGene00233064" = "Bm12803", "bmalayi-WBGene00223435" = "Bma-cht-1.2", "bmalayi-WBGene00224373" = "Bm4112", "bmalayi-WBGene00233954" = "Bm13693", "bmalayi-WBGene00223499" = "Bma-nmad-1", "bmalayi-WBGene00233055" = "Bm12794", "bmalayi-WBGene00225966" = "Bm5705", "bmalayi-WBGene00230858" = "Bma-hsp-17", "bmalayi-WBGene00225845" = "Bma-enpl-1", "bmalayi-WBGene00222412" = "Bm2151", "bmalayi-WBGene00223487" = "Bma-prdx-2", "bmalayi-WBGene00223805" = "Bma-rpt-4", "bmalayi-WBGene00224389" = "Bma-ned-8", "bmalayi-WBGene00231332" = "Bm11071", "bmalayi-WBGene00223115" = "Bm2854", "bmalayi-WBGene00226682" = "Bma-col-109", "bmalayi-WBGene00230319" = "Bma-pqn-35", "bmalayi-WBGene00222728" = "Bm2467", "bmalayi-WBGene00224893" = "Bma-ctg-1")) + xlab("Genes") + ylab("Cluster")
text_genes_clus_dotplot

ggsave(file = "Figure_3A.pdf", plot = text_genes_clus_dotplot, width = 15, height =5, dpi = 300)

```

```{r plot clus violins}

#Figure 3B-F

genes_of_interest <- c("bmalayi-WBGene00228202", "bmalayi-WBGene00223012", "bmalayi-WBGene00223435", "bmalayi-WBGene00233055", "bmalayi-WBGene00231332", "bmalayi-WBGene00223115", "bmalayi-WBGene00230858")

vln_list <- lapply(genes_of_interest, function(gene) {
  vln_plot <- VlnPlot(new_clus_names, assay = "SCT", features = gene, pt.size = 0, cols = fill, group.by = "seurat_clusters") + xlab("Cluster") + ylab("UMI counts per spot (normalized)")
  return(vln_plot)
})

vln_list

ggsave(file = paste0("Figure_3B_WBGene00228202_Bm7941", ".pdf"), plot = vln_list[[1]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3C_WBGene00223012_Bm2751", ".pdf"), plot = vln_list[[2]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3D_WBGene00223435_Bma-cht-1.2", ".pdf"), plot = vln_list[[3]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3E_WBGene00233055_Bm12794", ".pdf"), plot = vln_list[[4]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3F_WBGene00231332_Bm11071", ".pdf"), plot = vln_list[[5]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3F_WBGene00223115_Bm2854", ".pdf"), plot = vln_list[[6]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_3F_WBGene00230858_Bma-hsp-17", ".pdf"), plot = vln_list[[7]], width = 15, height =10, dpi = 300)

```
### Figure 4

```{r gly genes}

gly_genes <- c("WBGene00225841","WBGene00225960","WBGene00223600")
#glycolysis_genes

gly_genes_rename <- paste0("bmalayi-", gly_genes)
#glycolysis_genes_rename

```

```{r plot gly genes dotplot}

#Figure 4A

gly_text_dotplot <- DotPlot(new_clus_names, assay = "SCT", gly_genes_rename, cols = c("#fef0d9", "#b30000"), group.by = "seurat_clusters", dot.scale = 10) + RotatedAxis() + scale_x_discrete(labels = c("bmalayi-WBGene00225841" = "Bma-aldo-1", "bmalayi-WBGene00225960" = "Bm5699", "bmalayi-WBGene00223600" = "Bma-ldh-1")) + xlab("Genes") + ylab("Cluster")
gly_text_dotplot

ggsave(file = "Figure_4A.pdf", plot = gly_text_dotplot, width = 5, height =5, dpi = 300)
```

```{r plot gly violins}

#Figure 4B-D

gly_vln_list <- lapply(gly_genes_rename, function(gly_gene) {
  vln_plot <- VlnPlot(new_clus_names, assay = "SCT", features = gly_gene, pt.size = 0, cols = fill, group.by = "seurat_clusters") + xlab("Cluster") + ylab("UMI counts per spot (normalized)")
  return(vln_plot)
})

gly_vln_list

ggsave(file = paste0("Figure_4B_WBGene00225841_Bma-aldo-1", ".pdf"), plot = gly_vln_list[[1]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_4C_WBGene00225960_Bm5699", ".pdf"), plot = gly_vln_list[[2]], width = 15, height =10, dpi = 300)
ggsave(file = paste0("Figure_4D_WBGene00223600_Bma-ldh-1", ".pdf"), plot = gly_vln_list[[3]], width = 15, height =10, dpi = 300)

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

The outputs are the dimensionally reduced, integrated with harmony rds object with clusters, the deferentially expressed marker genes per cluster in an excel file in Supplementary Data 2, and the plots used in Figure 2B, Figure 3A-F, and Figure 4A-D. The dimensionally reduced, integrated with harmony, and clustered rds object of brugia count matrix and metadata is used to make the plots shown in Figure 2D and used for the functional term enrichment analysis performed to generate Figure 2E and Supplementary Data 3. The output clustered object is used in the analysis scripts 7_co_localization_analysis.Rmd, 8_genes_on_tissue_visualization.Rmd, and 9_wolbachia_abundance_variation.Rmd.
