---
title: "Co-localization analysis"
author: "Hailey Sounart"
date: "8/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(hdf5r)
  library(imager)
  library(akima)
  library(plotly)
  library(raster)
  library(dplyr)
  library(magrittr)
  library(ggplot2)
  library(imager) 
  library(magrittr)
  library(RColorBrewer)
  library(stringr)
  library(openxlsx)
  library(hash)
  library(readxl)
  library(purrr)
  library(harmony)
  library(DESeq2)
  library(performance)
  library(Hmisc)
  library(corrplot)
  library(pheatmap)
  library(performance)
  library(cowplot)
  library(ggpubr)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")
``` 

## Load normalized object

Use filtered, normalized object.
```{r read in normalized obj}

brugia_norm_obj <- readRDS("bmalay_filt_obj.list.nfeat.merged.rds")
brugia_norm_obj

brugia_norm_obj@meta.data

```


## Add wolbachia metadata to brugia object
```{r make wBm metadata}

brugia_norm_obj@meta.data$wolbachia <- ifelse((brugia_norm_obj@meta.data$counts.wBm >= 1), 1, 0)
brugia_norm_obj@meta.data

```
```{r check metadata}

identical(brugia_norm_obj@meta.data$wolbachia, brugia_norm_obj@meta.data$wolbachia)

```


## DE genes wolbachia to non-wolbachia spots

```{r calc co-local de genes}

de_tests <- c("wilcox","bimod","roc","t","negbinom","poisson","LR","MAST","DESeq2")

brugia_wBm_de_genes_lists <- lapply(de_tests, function(i) {
  FindMarkers(brugia_norm_obj, assay = "SCT", ident.1 = 1, ident.2 = 0, group.by = "wolbachia", logfc.threshold = 0.1, test.use = i)
})
brugia_wBm_de_genes_lists

```

## Keep wilcox and DESeq2 test DE genes

adj p values < 0.05 & avglogFC >= 0.1

```{r keep de genes}


brugia_wBm_de_genes_wilcox_p0.05_fc0.1 <- brugia_wBm_de_genes_lists[[1]][which(brugia_wBm_de_genes_lists[[1]]$p_val_adj < 0.05),]
brugia_wBm_de_genes_wilcox_p0.05_fc0.1

brugia_wBm_de_genes_deseq2_p0.05_fc0.1 <- brugia_wBm_de_genes_lists[[9]][which(brugia_wBm_de_genes_lists[[9]]$p_val_adj < 0.05),]
brugia_wBm_de_genes_deseq2_p0.05_fc0.1

```

## Combine lists (wilcox & DESeq2) and add annotations

Put wilcox and deseq2 infront of columns names
```{r combine de genes 1}

colnames(brugia_wBm_de_genes_wilcox_p0.05_fc0.1) <- paste0('seurat_wilcox_', colnames(brugia_wBm_de_genes_wilcox_p0.05_fc0.1))
brugia_wBm_de_genes_wilcox_p0.05_fc0.1

colnames(brugia_wBm_de_genes_deseq2_p0.05_fc0.1) <-  paste0('seurat_deseq2_', colnames(brugia_wBm_de_genes_deseq2_p0.05_fc0.1))
brugia_wBm_de_genes_deseq2_p0.05_fc0.1

``` 

Combine lists (wilcox & DESeq2)
```{r combine de genes 2}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb <- transform(merge(brugia_wBm_de_genes_wilcox_p0.05_fc0.1, brugia_wBm_de_genes_deseq2_p0.05_fc0.1,by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb


``` 

Rename rows/genes
``` {r combine de genes 3}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed <- brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_genes <- gsub("bmalayi-", "", rownames(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed))
rownames(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed) <- brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_genes
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed


```

Read in annotations
```{r add annotations}

#read in the bmalayi annotation file as a table
all_bmalayi_gene_annotations <- read.table("brugia_malayi.gene_biotype_product.map.tsv", sep="\t", header=T, stringsAsFactors = F)
all_bmalayi_gene_annotations

```

Add annotations into table
```{r add annotations 2}

#all_bmalayi_gene_annotations$gene_id
all_bmalayi_gene_annotations_table <- all_bmalayi_gene_annotations
rownames(all_bmalayi_gene_annotations_table) <- all_bmalayi_gene_annotations$gene_id
all_bmalayi_gene_annotations_table

all_bmalayi_gene_annotations_table <- all_bmalayi_gene_annotations_table[ , -which(names(all_bmalayi_gene_annotations_table) %in% "gene_id")]
all_bmalayi_gene_annotations_table

```

Merge annotations into DE gene table
```{r add annotations 3}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot <- transform(merge(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed, all_bmalayi_gene_annotations_table,by="row.names",all.x=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot


``` 

## Save the DE gene lists
```{r save colocal de genes}

saveRDS(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot, file = "brugia_colocal_wilcox_deseq2_p0.05_fc0.1_de_genes.rds")
write.xlsx(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot, file = "brugia_colocal_wilcox_deseq2_p0.05_fc0.1_de_genes.xlsx", col.names=TRUE, row.names=TRUE)

``` 

Output is the deferentially expressed brugia genes in Wolbachia+ verus Wolbachia- spots in an excel file and as an rds object, and is shown in Supplementary Table 5. The output is used to generate the plots for Figure 4A-D. 