---
title: "Co-localization analysis"
author: "Hailey Sounart"
date: "8/16/2022"
output: html_document
---

```{r, include = TRUE, echo = TRUE, message = TRUE}

start_time <- Sys.time()
start_time
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, warning =FALSE)
```

```{r load packages}
  library(STutility)
  library(ggplot2)
  library(openxlsx)
  library(EnhancedVolcano)
```

```{r set working directory}

setwd("<path to working directory where count matrices for all samples reside>")

``` 

## Load normalized object

Use filtered, normalized object.
```{r read in normalized obj}

brugia_norm_obj <- readRDS("bmalay_filt_obj.list.nfeat.merged.rds")
brugia_norm_obj

brugia_norm_obj@meta.data

```


## Add wolbachia metadata to brugia object
```{r make wBm metadata}

brugia_norm_obj@meta.data$wolbachia <- ifelse((brugia_norm_obj@meta.data$counts.wBm >= 1), 1, 0)
brugia_norm_obj@meta.data

```
```{r check metadata}

identical(brugia_norm_obj@meta.data$wolbachia, brugia_norm_obj@meta.data$wolbachia)

```


## DE genes wolbachia to non-wolbachia spots

```{r calc co-local de genes}

de_tests <- c("wilcox","bimod","roc","t","negbinom","poisson","LR","MAST","DESeq2")

brugia_wBm_de_genes_lists <- lapply(de_tests, function(i) {
  FindMarkers(brugia_norm_obj, assay = "SCT", ident.1 = 1, ident.2 = 0, group.by = "wolbachia", logfc.threshold = 0.1, test.use = i)
})
brugia_wBm_de_genes_lists

```

## Keep wilcox and DESeq2 test DE genes

adj p values < 0.05 & avglogFC >= 0.1

```{r keep de genes}


brugia_wBm_de_genes_wilcox_p0.05_fc0.1 <- brugia_wBm_de_genes_lists[[1]][which(brugia_wBm_de_genes_lists[[1]]$p_val_adj < 0.05),]
brugia_wBm_de_genes_wilcox_p0.05_fc0.1

brugia_wBm_de_genes_deseq2_p0.05_fc0.1 <- brugia_wBm_de_genes_lists[[9]][which(brugia_wBm_de_genes_lists[[9]]$p_val_adj < 0.05),]
brugia_wBm_de_genes_deseq2_p0.05_fc0.1

```

## Combine lists (wilcox & DESeq2) and add annotations

Put wilcox and deseq2 infront of columns names
```{r combine de genes 1}

colnames(brugia_wBm_de_genes_wilcox_p0.05_fc0.1) <- paste0('seurat_wilcox_', colnames(brugia_wBm_de_genes_wilcox_p0.05_fc0.1))
brugia_wBm_de_genes_wilcox_p0.05_fc0.1

colnames(brugia_wBm_de_genes_deseq2_p0.05_fc0.1) <-  paste0('seurat_deseq2_', colnames(brugia_wBm_de_genes_deseq2_p0.05_fc0.1))
brugia_wBm_de_genes_deseq2_p0.05_fc0.1

``` 

Combine lists (wilcox & DESeq2)
```{r combine de genes 2}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb <- transform(merge(brugia_wBm_de_genes_wilcox_p0.05_fc0.1, brugia_wBm_de_genes_deseq2_p0.05_fc0.1,by="row.names",all.x=TRUE,all.y=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb


``` 
Rename rows/genes
``` {r combine de genes 3}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed <- brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_genes <- gsub("bmalayi-", "", rownames(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed))
rownames(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed) <- brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_genes
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed


```

Read in annotations
```{r add annotations}

#read in the bmalayi annotation file as a table
all_bmalayi_gene_annotations <- read.table("brugia_malayi.gene_biotype_product.map.tsv", sep="\t", header=T, stringsAsFactors = F)
all_bmalayi_gene_annotations

```

Add annotations into table
```{r add annotations 2}

#all_bmalayi_gene_annotations$gene_id
all_bmalayi_gene_annotations_table <- all_bmalayi_gene_annotations
rownames(all_bmalayi_gene_annotations_table) <- all_bmalayi_gene_annotations$gene_id
all_bmalayi_gene_annotations_table

all_bmalayi_gene_annotations_table <- all_bmalayi_gene_annotations_table[ , -which(names(all_bmalayi_gene_annotations_table) %in% "gene_id")]
all_bmalayi_gene_annotations_table

```

Merge annotations into DE gene table
```{r add annotations 3}

brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot <- transform(merge(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed, all_bmalayi_gene_annotations_table,by="row.names",all.x=TRUE), row.names=Row.names, Row.names=NULL) #,all.y=TRUE
brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot


``` 
## Save the DE gene lists (Supplementary Data 5)
```{r save colocal de genes}

#Supplementary Data 5
saveRDS(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot, file = "Supplementary_Data_5.rds")
write.xlsx(brugia_wBm_de_genes_wilcox_deseq2_p0.05_fc0.1_comb_renamed_annot, file = "Supplementary_Data_5.xlsx", col.names=TRUE, row.names=TRUE)

``` 

## Plots (Figure 5A-B)

### Figure 5A

``` {r load clustered object}

brugia_clu_obj <- readRDS("bmalayi_clusters.rds")
brugia_clu_obj

```

```{r rename clus numbers}

#rename clusters

new_clus_names <- brugia_clu_obj
new_clus_names$seurat_clusters <- factor(brugia_clu_obj$seurat_clusters, 
         levels = c("0", "1", "2", "3"), 
         labels = c("1", "2", "3", "4"))

new_clus_names@meta.data
```
```{r make wBm metadata}

#add wolbachia column to metadata
new_clus_names@meta.data$wolbachia <- ifelse((new_clus_names@meta.data$counts.wBm >= 1), 1, 0)
new_clus_names@meta.data

```

```{r plot colocal umap}

#Figure 5A

wol_plot <- DimPlot(new_clus_names, label = FALSE, cols = c("#00FF7F","#FA8072"), pt.size = 5, group.by = "wolbachia", reduction = "umap")
wol_plot

ggsave(file = paste0("Figure_5A.pdf"), plot = wol_plot, width = 15, height =10, dpi = 300)
```
### Figure 5B

```{r get colocal de genes}

colocal_de_genes <- FindMarkers(brugia_norm_obj, assay = "SCT", ident.1 = 1, ident.2 = 0, group.by = "wolbachia", logfc.threshold = 0, min.cells.feature = 0, min.cells.group = 0, min.pct = 0)
```

```{r filt de genes}

#get top genes by fold change and significance 

sig_up_colocal_de_genes <- rownames(colocal_de_genes[(colocal_de_genes$p_val_adj < 0.05 & colocal_de_genes$avg_logFC > 0.4),])
#sig_up_colocal_de_genes
sig_down_colocal_de_genes <- rownames(colocal_de_genes[(colocal_de_genes$p_val_adj < 0.05 & colocal_de_genes$avg_logFC < -0.4),])
#sig_down_colocal_de_genes

sig_up_down_colocal_de_genes <- c(sig_up_colocal_de_genes, sig_down_colocal_de_genes)
#sig_up_down_colocal_de_genes

sig_up_down_colocal_de_genes_plus_text_genes <- c(sig_up_down_colocal_de_genes, "bmalayi-WBGene00224389", "bmalayi-WBGene00222232")
#sig_up_down_colocal_de_genes_plus_text_genes
```


```{r plot volcano plot}

#Figure 5B

coloca_volcano <- EnhancedVolcano(colocal_de_genes, rownames(colocal_de_genes), "avg_logFC", "p_val_adj", FCcutoff = 0, pCutoff = 0.05, labSize = 1.5, pointSize = 2, col = c("grey75", "grey75", "grey75", "#EE8262"), drawConnectors = TRUE, widthConnectors = 0.1, typeConnectors = "closed", endsConnectors = "first", lengthConnectors = unit(0.000001, "npc"), colConnectors = "grey10", selectLab = sig_up_down_colocal_de_genes_plus_text_genes, colAlpha = 1, labCol = "black", gridlines.minor = FALSE, gridlines.major = FALSE, xlim = c(min(colocal_de_genes[["avg_logFC"]], na.rm = TRUE) - 0.25, max(colocal_de_genes[["avg_logFC"]], na.rm = TRUE) + 0.25)) + xlab("Average log fold change")
coloca_volcano


coloca_volcano_2 <- EnhancedVolcano(colocal_de_genes, rownames(colocal_de_genes), "avg_logFC", "p_val_adj", FCcutoff = 0, pCutoff = 0.05, labSize = 1.5, pointSize = 2, col = c("grey75", "grey75", "grey75", "#008B45"), drawConnectors = TRUE, widthConnectors = 0.1, typeConnectors = "closed", endsConnectors = "first", lengthConnectors = unit(0.000001, "npc"), colConnectors = "grey10", selectLab = sig_up_down_colocal_de_genes_plus_text_genes, colAlpha = 1, labCol = "black", gridlines.minor = FALSE, gridlines.major = FALSE, xlim = c(min(colocal_de_genes[["avg_logFC"]], na.rm = TRUE) - 0.25, max(colocal_de_genes[["avg_logFC"]], na.rm = TRUE) + 0.25)) + xlab("Average log fold change")
coloca_volcano_2

ggsave(file = paste0("Figure_5B_1.pdf"), plot = coloca_volcano, dpi = 300, width = 10, height =8)
ggsave(file = paste0("Figure_5B_2.pdf"), plot = coloca_volcano_2, dpi = 300, width = 10, height =8)

```

```{r, include = TRUE, echo = TRUE, message = TRUE}

end_time <- Sys.time()
end_time
```

```{r, include = TRUE, echo = TRUE, message = TRUE}

elapsed_time <- round((end_time-start_time), 3)
elapsed_time
```

Output is the deferentially expressed brugia genes in Wolbachia+ verus Wolbachia- spots in an excel file and as an rds object, and is shown in Supplementary Data 5. The output plots are used in Figures 5A-B. 
